---
title: "NEON_group_project"
author: "AGB"
date: "2025-04-13"
output:
  word_document: 
    number_sections: true
  pdf_document: 
    toc: true
    number_sections: true
    toc_depth: 4
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Documents/Spring 2025/ECOL8322")
```

Plan of attack for document (Anna)
- Going to use longleaf pine
- Calculate standing carbon stocks for aboveground like assignment --- Completed
- Calculate standing carbon stocks for soil like assignment --- Completed
- Calculate standing carbon stocks for roots like assignment --- To be done
- Need to brainstorm for nitrogen how-to --- To be done

Environment
```{r}											
#install.packages("neonUtilities")						
library(neonUtilities)								
library(tidyverse)
library(lubridate)
```

#DSNY -- Tree Carbon Stocks
Downloading the data of interest from NEON
```{r}												
DSNY.trees <- loadByProduct(dpID="DP1.10098.001",
site=c("DSNY"),
startdate="2012-01", enddate="2024-12",
package="basic")

list2env(DSNY.trees, .GlobalEnv)
```	

Write CSVs for four tables extracted
```{r}											
#readme_10098
write.csv(readme_10098, "readme_10098_DSNY.csv")

#vst_perplotperyear
write.csv(vst_perplotperyear, "vst_perplotperyear_DSNY.csv")

#vst_apparentindividual
write.csv(vst_apparentindividual, "vst_apparentindividual_DSNY.csv")

#variables_10098
write.csv(variables_10098, "variables_10098_DSNY.csv")
```

readme_10098 has a description of all the dataframes and variables_10098 defines all the variables in each dataframe. As you are working with NEON data in the future, you’ll want to familiarize yourself with these, but for the rest of this activity we will not need to refer to them.

Read in one of the CSVs we wrote above and revise to be compatible with the one that holds tree measurements
```{r}											
vst_perplotperyear_DSNY <- read.csv("vst_perplotperyear_DSNY.csv")
variables_10098_DSNY <- read.csv("variables_10098_DSNY.csv")

#making a new dataframe with selected variables
plot.info.vst.DSNY <- vst_perplotperyear_DSNY %>%
select(date, domainID, siteID, plotID, plotType, nlcdClass, totalSampledAreaTrees)

#making a new variable that containers only year
plot.info.vst.DSNY <- plot.info.vst.DSNY %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each PLOT
plot.info.vst.DSNY.rec <- plot.info.vst.DSNY %>%
group_by(plotID)%>%
filter(year==max(year))

#remove NAs from the totalSampledAreaTrees variable 
plot.info.vst.DSNY.rec <- plot.info.vst.DSNY.rec %>% filter(totalSampledAreaTrees!="NA")

```

Repeating the process above for individual dataframe
```{r}											
vst_apparentindividual_DSNY <- read.csv("vst_apparentindividual_DSNY.csv")

#making a new dataframe with selected variables
individuals.vst.DSNY <- vst_apparentindividual_DSNY %>%
select(date, siteID, plotID, individualID, growthForm, stemDiameter, plantStatus)

#making a new variable that containers only year
individuals.vst.DSNY <- individuals.vst.DSNY %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each INDIVIDUAL
most.recent.individuals.vst.DSNY <- individuals.vst.DSNY %>%
group_by(individualID) %>%
filter(year==max(year))

#making new dataframe where you filter the prior dataframe for only live trees
most.recent.live.plants.DSNY.vst <- most.recent.individuals.vst.DSNY %>% filter(str_detect(plantStatus, "Live")) #little unsure here

#making new dataframe for single bole trees, multi-bole trees, and small trees only
most.recent.live.trees1.DSNY <- most.recent.live.plants.DSNY.vst %>%
  filter(growthForm=="single bole tree" | growthForm=="multi-bole tree" | growthForm=="small tree")

#for dataframe above now we are excluding trees that have a recorded stem diameter of NA, we need to do this for analysis
most.recent.live.trees.DSNY <- subset(most.recent.live.trees1.DSNY, stemDiameter>0)

```

Joining dataframes for DSNY
```{r}

# str(most.recent.live.trees.DSNY)
# str(plot.info.vst.DSNY.rec)

tree.data.DSNY <- left_join(most.recent.live.trees.DSNY, plot.info.vst.DSNY.rec)
# str(tree.data.DSNY) #XX

tree.data.DSNY.complete<-subset(tree.data.DSNY, totalSampledAreaTrees>0)
# str(tree.data.DSNY.complete) #XX

#We lost information for XX trees when joining dataframes together.

```

```{r}

#need a new column in joined dataframe that represents biomass of each tree (in kg) called AGB
#mixed hardwood allometric equation from jenkins et al 2003
#Equation that I need to use is ABG (kg) = exp(B0 + B1 x ln(dbh))
tree.data.DSNY.complete$AGB <- exp(-2.4800 + (2.4835*log(tree.data.DSNY.complete$stemDiameter))) #this equation gives kg

#now need to create a new dataframe called AGB.trees.by.plot.DSNY
AGB.trees.by.plot.DSNY <- tree.data.DSNY.complete %>%
  group_by(siteID, plotID, totalSampledAreaTrees) %>% #grouping variables together
  summarise(plot.AGB = sum(AGB)) #summing by grouping variables 
# View(AGB.trees.by.plot.DSNY)

AGB.trees.by.plot.DSNY$AGB.kg.m2<-AGB.trees.by.plot.DSNY$plot.AGB/AGB.trees.by.plot.DSNY$totalSampledAreaTrees #making it as kg/m2

#Okay, need to get m2 to ha (10,000 m2 = 1 ha)
#then to get kg to mg (1,000 kg = 1 mg)

AGB.trees.by.plot.DSNY <- AGB.trees.by.plot.DSNY %>%
  mutate(AGB.kg.m2 = plot.AGB/totalSampledAreaTrees) %>%
  mutate(AGB.Mg.ha = (((AGB.kg.m2)*10))) %>%
  mutate(AGB.C.Mg.ha = AGB.Mg.ha*0.47)
# View(AGB.trees.by.plot.DSNY)


```

```{r}
ggplot(data=AGB.trees.by.plot.DSNY, aes(x=plotID, y=AGB.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```


#JERC -- Tree Carbon Stocks
Downloading the data of interest from NEON
```{r}												
JERC.trees <- loadByProduct(dpID="DP1.10098.001", 	
site=c("JERC"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(JERC.trees, .GlobalEnv)
```	

Write CSVs for four tables extracted
```{r}											
#readme_10098
write.csv(readme_10098, "readme_10098_JERC.csv")

#vst_perplotperyear
write.csv(vst_perplotperyear, "vst_perplotperyear_JERC.csv")

#vst_apparentindividual
write.csv(vst_apparentindividual, "vst_apparentindividual_JERC.csv")

#variables_10098
write.csv(variables_10098, "variables_10098_JERC.csv")
```

readme_10098 has a description of all the dataframes and variables_10098 defines all the variables in each dataframe. As you are working with NEON data in the future, you’ll want to familiarize yourself with these, but for the rest of this activity we will not need to refer to them.

Read in one of the CSVs we wrote above and revise to be compatible with the one that holds tree measurements
```{r}											
vst_perplotperyear_JERC <- read.csv("vst_perplotperyear_JERC.csv")
variables_10098_JERC <- read.csv("variables_10098_JERC.csv")

#making a new dataframe with selected variables
plot.info.vst.JERC <- vst_perplotperyear_JERC %>%
select(date, domainID, siteID, plotID, plotType, nlcdClass, totalSampledAreaTrees)

#making a new variable that containers only year
plot.info.vst.JERC <- plot.info.vst.JERC %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each PLOT
plot.info.vst.JERC.rec <- plot.info.vst.JERC %>%
group_by(plotID)%>%
filter(year==max(year))

#remove NAs from the totalSampledAreaTrees variable 
plot.info.vst.JERC.rec <- plot.info.vst.JERC.rec %>% filter(totalSampledAreaTrees!="NA")

```

Repeating the process above for individual dataframe
```{r}											
vst_apparentindividual_JERC <- read.csv("vst_apparentindividual_JERC.csv")

#making a new dataframe with selected variables
individuals.vst.JERC <- vst_apparentindividual_JERC %>%
select(date, siteID, plotID, individualID, growthForm, stemDiameter, plantStatus)

#making a new variable that containers only year
individuals.vst.JERC <- individuals.vst.JERC %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each INDIVIDUAL
most.recent.individuals.vst.JERC <- individuals.vst.JERC %>%
group_by(individualID) %>%
filter(year==max(year))

#making new dataframe where you filter the prior dataframe for only live trees
most.recent.live.plants.JERC.vst <- most.recent.individuals.vst.JERC %>%
filter(str_detect(plantStatus, "Live")) #little unsure here

#making new dataframe for single bole trees, multi-bole trees, and small trees only
most.recent.live.trees1.JERC <- most.recent.live.plants.JERC.vst %>%
  filter(growthForm=="single bole tree" | growthForm=="multi-bole tree" | growthForm=="small tree")

#for dataframe above now we are excluding trees that have a recorded stem diameter of NA, we need to do this for analysis
most.recent.live.trees.JERC <- subset(most.recent.live.trees1.JERC, stemDiameter>0)

```

Joining dataframes for JERC
```{r}

# str(most.recent.live.trees.JERC)
# str(plot.info.vst.JERC.rec)

tree.data.JERC <- left_join(most.recent.live.trees.JERC, plot.info.vst.JERC.rec)
# str(tree.data.JERC) #XXX

tree.data.JERC.complete<-subset(tree.data.JERC, totalSampledAreaTrees>0)
# str(tree.data.JERC.complete) #XXX

#We lost information for XXX trees when joining dataframes together.

```

```{r}

#need a new column in joined dataframe that represents biomass of each tree (in kg) called AGB
#mixed hardwood allometric equation from jenkins et al 2003
#Equation that I need to use is ABG (kg) = exp(B0 + B1 x ln(dbh))
tree.data.JERC.complete$AGB <- exp(-2.4800 + (2.4835*log(tree.data.JERC.complete$stemDiameter))) #this equation gives kg

#now need to create a new dataframe called AGB.trees.by.plot.JERC
AGB.trees.by.plot.JERC <- tree.data.JERC.complete %>%
  group_by(siteID, plotID, totalSampledAreaTrees) %>% #grouping variables together
  summarise(plot.AGB = sum(AGB)) #summing by grouping variables 
View(AGB.trees.by.plot.JERC)

AGB.trees.by.plot.JERC$AGB.kg.m2<-AGB.trees.by.plot.JERC$plot.AGB/AGB.trees.by.plot.JERC$totalSampledAreaTrees #making it as kg/m2

#Okay, need to get m2 to ha (10,000 m2 = 1 ha)
#then to get kg to mg (1,000 kg = 1 mg)

AGB.trees.by.plot.JERC <- AGB.trees.by.plot.JERC %>%
  mutate(AGB.kg.m2 = plot.AGB/totalSampledAreaTrees) %>%
  mutate(AGB.Mg.ha = (((AGB.kg.m2)*10))) %>%
  mutate(AGB.C.Mg.ha = AGB.Mg.ha*0.47)
# View(AGB.trees.by.plot.JERC)

```

```{r}
ggplot(data=AGB.trees.by.plot.JERC, aes(x=plotID, y=AGB.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#OSBS -- Tree Carbon Stocks
Downloading the data of interest from NEON
```{r}												
OSBS.trees <- loadByProduct(dpID="DP1.10098.001", 	
site=c("OSBS"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(OSBS.trees, .GlobalEnv)
```	

Write CSVs for four tables extracted
```{r}											
#readme_10098
write.csv(readme_10098, "readme_10098_OSBS.csv")

#vst_perplotperyear
write.csv(vst_perplotperyear, "vst_perplotperyear_OSBS.csv")

#vst_apparentindividual
 write.csv(vst_apparentindividual, "vst_apparentindividual_OSBS.csv")

#variables_10098
write.csv(variables_10098, "variables_10098_OSBS.csv")
```

readme_10098 has a description of all the dataframes and variables_10098 defines all the variables in each dataframe. As you are working with NEON data in the future, you’ll want to familiarize yourself with these, but for the rest of this activity we will not need to refer to them.

Read in one of the CSVs we wrote above and revise to be compatible with the one that holds tree measurements
```{r}											
vst_perplotperyear_OSBS <- read.csv("vst_perplotperyear_OSBS.csv")
variables_10098_OSBS <- read.csv("variables_10098_OSBS.csv")

#making a new dataframe with selected variables
plot.info.vst.OSBS <- vst_perplotperyear_OSBS %>%
select(date, domainID, siteID, plotID, plotType, nlcdClass, totalSampledAreaTrees)

#making a new variable that containers only year
plot.info.vst.OSBS <- plot.info.vst.OSBS %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each PLOT
plot.info.vst.OSBS.rec <- plot.info.vst.OSBS %>%
group_by(plotID)%>%
filter(year==max(year))

#remove NAs from the totalSampledAreaTrees variable 
plot.info.vst.OSBS.rec <- plot.info.vst.OSBS.rec %>% filter(totalSampledAreaTrees!="NA")

```

Repeating the process above for individual dataframe
```{r}											
vst_apparentindividual_OSBS <- read.csv("vst_apparentindividual_OSBS.csv")

#making a new dataframe with selected variables
individuals.vst.OSBS <- vst_apparentindividual_OSBS %>%
select(date, siteID, plotID, individualID, growthForm, stemDiameter, plantStatus)

#making a new variable that containers only year
individuals.vst.OSBS <- individuals.vst.OSBS %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each INDIVIDUAL
most.recent.individuals.vst.OSBS <- individuals.vst.OSBS %>%
group_by(individualID) %>%
filter(year==max(year))

#making new dataframe where you filter the prior dataframe for only live trees
most.recent.live.plants.OSBS.vst <- most.recent.individuals.vst.OSBS %>%
filter(str_detect(plantStatus, "Live")) #little unsure here

#making new dataframe for single bole trees, multi-bole trees, and small trees only
most.recent.live.trees1.OSBS <- most.recent.live.plants.OSBS.vst %>%
  filter(growthForm=="single bole tree" | growthForm=="multi-bole tree" | growthForm=="small tree")

#for dataframe above now we are excluding trees that have a recorded stem diameter of NA, we need to do this for analysis
most.recent.live.trees.OSBS <- subset(most.recent.live.trees1.OSBS, stemDiameter>0)

```

Joining dataframes for OSBS
```{r}

# str(most.recent.live.trees.OSBS)
# str(plot.info.vst.OSBS.rec)

tree.data.OSBS <- left_join(most.recent.live.trees.OSBS, plot.info.vst.OSBS.rec)
# str(tree.data.OSBS) #XXX

tree.data.OSBS.complete<-subset(tree.data.OSBS, totalSampledAreaTrees>0)
# str(tree.data.OSBS.complete) #XXX

#We lost information for XXX trees when joining dataframes together.

```

```{r}

#need a new column in joined dataframe that represents biomass of each tree (in kg) called AGB
#mixed hardwood allometric equation from jenkins et al 2003
#Equation that I need to use is ABG (kg) = exp(B0 + B1 x ln(dbh))
tree.data.OSBS.complete$AGB <- exp(-2.4800 + (2.4835*log(tree.data.OSBS.complete$stemDiameter))) #this equation gives kg

#now need to create a new dataframe called AGB.trees.by.plot.OSBS
AGB.trees.by.plot.OSBS <- tree.data.OSBS.complete %>%
  group_by(siteID, plotID, totalSampledAreaTrees) %>% #grouping variables together
  summarise(plot.AGB = sum(AGB)) #summing by grouping variables 
# View(AGB.trees.by.plot.OSBS)

AGB.trees.by.plot.OSBS$AGB.kg.m2<-AGB.trees.by.plot.OSBS$plot.AGB/AGB.trees.by.plot.OSBS$totalSampledAreaTrees #making it as kg/m2

#Okay, need to get m2 to ha (10,000 m2 = 1 ha)
#then to get kg to mg (1,000 kg = 1 mg)

AGB.trees.by.plot.OSBS <- AGB.trees.by.plot.OSBS %>%
  mutate(AGB.kg.m2 = plot.AGB/totalSampledAreaTrees) %>%
  mutate(AGB.Mg.ha = (((AGB.kg.m2)*10))) %>%
  mutate(AGB.C.Mg.ha = AGB.Mg.ha*0.47)
# View(AGB.trees.by.plot.OSBS)

```

```{r}
ggplot(data=AGB.trees.by.plot.OSBS, aes(x=plotID, y=AGB.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```



#Joining dataframes across D03 (Southeastern longleaf pine forests) for Tree Carbon Stocks
```{r}

# str(AGB.trees.by.plot.DSNY) #XX
# str(AGB.trees.by.plot.JERC) #XX
# str(AGB.trees.by.plot.OSBS) #XX

AGB.trees.by.plot.D03 <- full_join(AGB.trees.by.plot.DSNY, AGB.trees.by.plot.JERC)
AGB.trees.by.plot.D03 <- full_join(AGB.trees.by.plot.D03, AGB.trees.by.plot.OSBS)
#str(AGB.trees.by.plot.D03)
```

#Visualizing Tree Carbon Stocks by site of D03
```{r}

AGB.trees.by.plot.D03.summarise <- AGB.trees.by.plot.D03 %>%
 group_by(siteID) %>%
 summarise(AGB.C.Mg.ha_siteAverage = mean(AGB.C.Mg.ha),
           AGB.C.Mg.ha_siteSD = sd(AGB.C.Mg.ha))
 
ggplot(data=AGB.trees.by.plot.D03, aes(x=siteID, y=AGB.C.Mg.ha, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = AGB.trees.by.plot.D03.summarise, aes(x = siteID, y = AGB.C.Mg.ha_siteAverage, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = AGB.trees.by.plot.D03.summarise, aes(x= siteID, y = AGB.C.Mg.ha_siteAverage, ymin = (AGB.C.Mg.ha_siteAverage-AGB.C.Mg.ha_siteSD), ymax = (AGB.C.Mg.ha_siteAverage+AGB.C.Mg.ha_siteSD), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Tree Carbon Stocks by Site of D03") +
        ylab("Carbon Stock (Mg/ha)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none") 

```


#DSNY -- Root Carbon Stocks
Create a chunk where you extract root biomass using NEON utilies package, get into environment, and write csvs into the environment
```{r}

DSNY.roots <- loadByProduct(dpID="DP1.10067.001", 	
site=c("DSNY"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(DSNY.roots, .GlobalEnv)

#readme_10067
write.csv(readme_10067, "readme_10067_DSNY.csv")

#bbc_percore
write.csv(bbc_percore, "bbc_percore_DSNY.csv")

#bbc_rootmass
write.csv(bbc_rootmass, "bbc_rootmass_DSNY.csv")

#variables_10067
write.csv(variables_10067, "variables_10067_DSNY.csv")

```

Create chunk selecting variables to create new dfs
```{r}

bbc_rootmass_DSNY <- read.csv("bbc_rootmass_DSNY.csv")
bbc_percore_DSNY <- read.csv("bbc_percore_DSNY.csv")

root.mass.DSNY <- bbc_rootmass_DSNY %>%
select(domainID, plotID, sampleID, subsampleID, collectDate, sizeCategory, rootStatus, dryMass)

root.core.DSNY <- bbc_percore_DSNY %>%
 select(domainID, plotID, subplotID, sampleID, clipID, coreID, collectDate, rootSampleArea, rootSampleDepth, coreDiameter)


root.data.DSNY <- left_join(root.mass.DSNY, root.core.DSNY)

root.data.DSNY$year <- substr(root.data.DSNY$collectDate, start=1, stop=4) 

```

Root calculations and conversions
```{r}

roots.by.sample.DSNY <- root.data.DSNY %>% 
  group_by(plotID, sampleID, rootSampleArea, rootSampleDepth) %>%
  summarise(total.mass = sum(dryMass))

roots.by.sample.DSNY.final <- roots.by.sample.DSNY %>%
  mutate(root.gC.sample = (total.mass * 0.47)) %>%
  mutate(root.gC.m2 = (root.gC.sample/rootSampleArea)) %>%
  mutate(root.MgC.ha = (root.gC.m2 * 0.01))
 
roots.by.plot.DSNY <- roots.by.sample.DSNY.final %>% 
  group_by(plotID) %>%
  summarise(mean.root.MgC.ha = mean(root.MgC.ha))

```

Visualize the data
```{r}

ggplot(data=roots.by.plot.DSNY, aes(x=plotID, y=mean.root.MgC.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#JERC -- Root Carbon Stocks
Create a chunk where you extract root biomass using NEON utilies package, get into environment, and write csvs into the environment
```{r}

JERC.roots <- loadByProduct(dpID="DP1.10067.001",
site=c("JERC"),
startdate="2012-01", enddate="2024-12",
package="basic")

list2env(JERC.roots, .GlobalEnv)

#readme_10067
write.csv(readme_10067, "readme_10067_JERC.csv")

#bbc_percore
write.csv(bbc_percore, "bbc_percore_JERC.csv")

#bbc_rootmass
write.csv(bbc_rootmass, "bbc_rootmass_JERC.csv")

#variables_10067
write.csv(variables_10067, "variables_10067_JERC.csv")

```

Create chunk selecting variables to create new dfs
```{r}

bbc_rootmass_JERC <- read.csv("bbc_rootmass_JERC.csv")
bbc_percore_JERC <- read.csv("bbc_percore_JERC.csv")

root.mass.JERC <- bbc_rootmass_JERC %>%
select(domainID, plotID, sampleID, subsampleID, collectDate, sizeCategory, rootStatus, dryMass)

root.core.JERC <- bbc_percore_JERC %>%
 select(domainID, plotID, subplotID, sampleID, clipID, coreID, collectDate, rootSampleArea, rootSampleDepth, coreDiameter)


root.data.JERC <- left_join(root.mass.JERC, root.core.JERC)

root.data.JERC$year <- substr(root.data.JERC$collectDate, start=1, stop=4) 

```

Root calculations and conversions
```{r}

roots.by.sample.JERC <- root.data.JERC %>% 
  group_by(plotID, sampleID, rootSampleArea, rootSampleDepth) %>%
  summarise(total.mass = sum(dryMass))

roots.by.sample.JERC.final <- roots.by.sample.JERC %>%
  mutate(root.gC.sample = (total.mass * 0.47)) %>%
  mutate(root.gC.m2 = (root.gC.sample/rootSampleArea)) %>%
  mutate(root.MgC.ha = (root.gC.m2 * 0.01))
 
roots.by.plot.JERC <- roots.by.sample.JERC.final %>% 
  group_by(plotID) %>%
  summarise(mean.root.MgC.ha = mean(root.MgC.ha))

```

Visualize the data
```{r}

ggplot(data=roots.by.plot.JERC, aes(x=plotID, y=mean.root.MgC.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#OSBS -- Root Carbon Stocks
Create a chunk where you extract root biomass using NEON utilies package, get into environment, and write csvs into the environment
```{r}

OSBS.roots <- loadByProduct(dpID="DP1.10067.001", 	
site=c("OSBS"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

 list2env(OSBS.roots, .GlobalEnv)

#readme_10067
write.csv(readme_10067, "readme_10067_OSBS.csv")

#bbc_percore
write.csv(bbc_percore, "bbc_percore_OSBS.csv")

#bbc_rootmass
write.csv(bbc_rootmass, "bbc_rootmass_OSBS.csv")

#variables_10067
write.csv(variables_10067, "variables_10067_OSBS.csv")

```

Create chunk selecting variables to create new dfs
```{r}

bbc_rootmass_OSBS <- read.csv("bbc_rootmass_OSBS.csv")
bbc_percore_OSBS <- read.csv("bbc_percore_OSBS.csv")

root.mass.OSBS <- bbc_rootmass_OSBS %>%
select(domainID, plotID, sampleID, subsampleID, collectDate, sizeCategory, rootStatus, dryMass)

root.core.OSBS <- bbc_percore_OSBS %>%
 select(domainID, plotID, subplotID, sampleID, clipID, coreID, collectDate, rootSampleArea, rootSampleDepth, coreDiameter)


root.data.OSBS <- left_join(root.mass.OSBS, root.core.OSBS)

root.data.OSBS$year <- substr(root.data.OSBS$collectDate, start=1, stop=4) 

```

Root calculations and conversions
```{r}

roots.by.sample.OSBS <- root.data.OSBS %>% 
  group_by(plotID, sampleID, rootSampleArea, rootSampleDepth) %>%
  summarise(total.mass = sum(dryMass))

roots.by.sample.OSBS.final <- roots.by.sample.OSBS %>%
  mutate(root.gC.sample = (total.mass * 0.47)) %>%
  mutate(root.gC.m2 = (root.gC.sample/rootSampleArea)) %>%
  mutate(root.MgC.ha = (root.gC.m2 * 0.01))
 
roots.by.plot.OSBS <- roots.by.sample.OSBS.final %>% 
  group_by(plotID) %>%
  summarise(mean.root.MgC.ha = mean(root.MgC.ha))

```

Visualize the data
```{r}

ggplot(data=roots.by.plot.OSBS, aes(x=plotID, y=mean.root.MgC.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#Joining dataframes across D03 (Southeastern longleaf pine forests) for Tree Carbon Stocks
```{r}

# str(roots.by.plot.DSNY) #30
# str(roots.by.plot.JERC) #20
# str(roots.by.plot.OSBS) #20

roots.by.plot.D03 <- full_join(roots.by.plot.DSNY, roots.by.plot.JERC)
roots.by.plot.D03 <- full_join(roots.by.plot.D03, roots.by.plot.OSBS)
#str(roots.by.plot.D03)
```

#Visualizing Tree Carbon Stocks by site of D03
```{r}

roots.by.plot.D03$siteID <- str_sub(roots.by.plot.D03$plotID, 1, 4)

roots.by.plot.D03.summarise <- roots.by.plot.D03 %>%
 group_by(siteID) %>%
 summarise(mean.root.MgC.ha_siteAverage = mean(mean.root.MgC.ha),
           mean.root.MgC.ha_siteSD = sd(mean.root.MgC.ha)) 
 
ggplot(data=roots.by.plot.D03, aes(x=siteID, y=mean.root.MgC.ha, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = roots.by.plot.D03.summarise, aes(x = siteID, y = mean.root.MgC.ha_siteAverage, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = roots.by.plot.D03.summarise, aes(x= siteID, y = mean.root.MgC.ha_siteAverage, ymin = (mean.root.MgC.ha_siteAverage-mean.root.MgC.ha_siteSD), ymax = (mean.root.MgC.ha_siteAverage+mean.root.MgC.ha_siteSD), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Root Carbon Stocks by Site of D03") +
        ylab("Carbon Stock (Mg/ha)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none") 

```
