---
title: "04.16.25_Bushong_NEON_group_project_Nitrogen"
author: "AGB"
date: "2025-04-16"
output: html_document
---

Set working directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Documents/Spring 2025/ECOL8322")
```

Environment
```{r}											
#install.packages("neonUtilities")						
library(neonUtilities)								
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(ggbreak)
library(patchwork)

#Necessary packages / commands for installling neonNTrans. Follow directions to update packages as needed. You can try to call for compilation for packages, but use binary if needed.
#library(devtools) 
#install_github("NEONScience/NEON-Nitrogen-Transformations/neonNTrans", #dependencies=TRUE)  
#library(neonNTrans) 

```

Downloading the data of interest from NEON, listing, and csvs
```{r}												

#Commented out code is relevant if you are able to use neonNTrans package in your R version, which I (Anna) cannot
      # DSNY.soil.nitro <- loadByProduct(dpID="DP1.10086.001", 	
      # site=c("DSNY","JERC","OSBS"), 		
      # startdate="2012-01", enddate="2024-12", 				
      # package="basic")
      
      #list2env(DSNY.soil.nitro, .GlobalEnv)
      
      #spc_biogeochem
      #write.csv(spc_biogeochem, "spc_biogeochem.csv")
      
      #spc_bulkdensity
      #write.csv(spc_bulkdensity, "spc_bulkdensity.csv")
      
      
      # test.out <- def.calc.ntrans(kclInt = DSNY.soil.nitro$ntr_internalLab, kclIntBlank = DSNY.soil.nitro$ntr_internalLabBlanks, kclExt = DSNY.soil.nitro$ntr_externalLab, soilMoist = DSNY.soil.nitro$sls_soilMoisture, dropConditions = c("extract stored at incorrect temperature", "soil stored at incorrect temperature", "mass uncertain", "volume uncertain"))


data.summary.D03 <- read.csv("datasummary.D03.csv")
str(data.summary.D03)


# Creating variable for siteID aand soil horizon s an independent variable in the dataframe by duplicating sample ID, subsetting the string, separating by the hyphen, and adding back to the original dataframe
data.summary.D03$siteID = substr(data.summary.D03$sampleID, 1,4)
data.summary.D03$siteID_tosplit = substr(data.summary.D03$sampleID, 1,10)
hold.name <- strsplit(data.summary.D03$siteID_tosplit, split="-")
hold.name.df <- as.data.frame(matrix(unlist(hold.name), ncol=2, byrow=T))
colnames(hold.name.df) <- c("plotID", "soilHorizon")
data.summary.D03$plotID <- hold.name.df$plotID
data.summary.D03$soilHorizon <- hold.name.df$soilHorizon

```


Notably, the Inorganic nitrogen metric provided through the neonNTrans package includes nitrite, which we are not as interested in, but some micorbes being able to utilize it as an intermediate. We could not a clearly identifiable way to isolate nitrate from nitrite with the available data.


# Figures at the plot-level by site
```{r}
data.summary.D03.DSNY <- data.summary.D03 %>%
  filter(siteID == "DSNY")
data.summary.D03.JERC <- data.summary.D03 %>%
  filter(siteID == "JERC")
data.summary.D03.OSBS <- data.summary.D03 %>%
  filter(siteID == "OSBS")

exploratory.InorgN.DSNY.plot <- ggplot(data=data.summary.D03.DSNY, aes(x=plotID, y=soilInorganicNugPerGram)) + 
  geom_point(size=1.5, alpha = 1)  +
        #scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Average Soil Inorganic N by Plot within DSNY of D03") +
       #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
        ylab("Inorganic N (mcg/g)") +
     geom_jitter() +
    theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
exploratory.InorgN.DSNY.plot

exploratory.InorgN.JERC.plot <- ggplot(data=data.summary.D03.JERC, aes(x=plotID, y=soilInorganicNugPerGram)) + 
  geom_point(size=1.5, alpha = 1)  +
        #scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Average Soil Inorganic N by Plot within JERC of D03") +
       #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
        ylab("Inorganic N (mcg/g)") +
     geom_jitter() +
    theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
exploratory.InorgN.JERC.plot

exploratory.InorgN.OSBSplot <- ggplot(data=data.summary.D03.OSBS, aes(x=plotID, y=soilInorganicNugPerGram)) + 
  geom_point(size=1.5, alpha = 1)  +
        #scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Average Soil Inorganic N by Plot within OSBS of D03") +
       #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
        ylab("Inorganic N (mcg/g)") +
     geom_jitter() +
    theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
exploratory.InorgN.OSBSplot


```

# Figures at the site-level
```{r}

# Initial visualization of data for Inorganic N
data.summary.D03.Exploratory <- data.summary.D03 %>%
 group_by(siteID) %>%
 summarise(soilInorganicNugPerGram_average = mean(soilInorganicNugPerGram),
           soilInorganicNugPerGram_sd = sd(soilInorganicNugPerGram))

exploratory.InorgN.D03 <- ggplot(data=data.summary.D03, aes(x=siteID, y=soilInorganicNugPerGram, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = data.summary.D03.Exploratory, aes(x = siteID, y = soilInorganicNugPerGram_average, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = data.summary.D03.Exploratory, aes(x= siteID, y = soilInorganicNugPerGram_average, ymin = (soilInorganicNugPerGram_average-soilInorganicNugPerGram_sd), ymax = (soilInorganicNugPerGram_average+soilInorganicNugPerGram_sd), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Average Soil Inorganic N by Site of D03") +
       scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
        ylab("Inorganic N (mcg/g)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none")

# Initial visualization of data for Ammonium
data.summary.D03.Exploratory.NH4 <- data.summary.D03 %>%
 group_by(siteID) %>%
 summarise(soilAmmoniumNugPerGram_average = mean(soilAmmoniumNugPerGram),
           soilAmmoniumNugPerGram_sd = sd(soilAmmoniumNugPerGram))

exploratory.ammonium.D03 <- ggplot(data=data.summary.D03, aes(x=siteID, y=soilAmmoniumNugPerGram, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = data.summary.D03.Exploratory.NH4, aes(x = siteID, y = soilAmmoniumNugPerGram_average, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = data.summary.D03.Exploratory.NH4, aes(x= siteID, y = soilAmmoniumNugPerGram_average, ymin = (soilAmmoniumNugPerGram_average-soilAmmoniumNugPerGram_sd), ymax = (soilAmmoniumNugPerGram_average+soilAmmoniumNugPerGram_sd), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Average Soil Ammonium by Site of D03") +
       scale_y_continuous(breaks = seq(-10, 80, 10), limits = c(-10, 65)) +
      scale_y_break(c(32, 48)) +
        ylab("Ammonium (mcg/g)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
              axis.text.y.right = element_blank(), #deals with weird ggbreak glitch
              axis.line.y.right = element_blank(), #deals with weird ggbreak glitch
              axis.ticks.y.right = element_blank(), #deals with weird ggbreak glitch
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none")

exploratory.InorgN.D03
exploratory.ammonium.D03

# Initial visualization of data for Mineralization
data.summary.D03.Exploratory.Mineralization <- data.summary.D03 %>%
 group_by(siteID) %>%
 summarise(netNminugPerGramPerDay_average = mean(netNminugPerGramPerDay), #averaging a rate, always messes with my head if appropriately statistically
           netNminugPerGramPerDay_sd = sd(netNminugPerGramPerDay))

exploratory.mineralization.D03 <- ggplot(data=data.summary.D03, aes(x=siteID, y=netNminugPerGramPerDay, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = data.summary.D03.Exploratory.Mineralization, aes(x = siteID, y = netNminugPerGramPerDay_average, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = data.summary.D03.Exploratory.Mineralization, aes(x= siteID, y = netNminugPerGramPerDay_average, ymin = (netNminugPerGramPerDay_average-netNminugPerGramPerDay_sd), ymax = (netNminugPerGramPerDay_average+netNminugPerGramPerDay_sd), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Average Soil N Mineralization rate by Site of D03") +
       #scale_y_continuous(breaks = seq(-10, 80, 10), limits = c(-10, 65)) +
     # scale_y_break(c(32, 48)) +
        ylab("Mineralization ((mcg/g)/day)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
             #axis.text.y.right = element_blank(), #deals with weird ggbreak glitch
              #axis.line.y.right = element_blank(), #deals with weird ggbreak glitch
              #axis.ticks.y.right = element_blank(), #deals with weird ggbreak glitch
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none")

exploratory.mineralization.D03 #note, getting a missing values warning
```

