---
title: "NEON_group_project"
author: "Andrea Rivera"
date: "2025-04-23"
output:
  word_document: 
    number_sections: true
  pdf_document: 
    toc: true
    number_sections: true
    toc_depth: 4
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/amriv/OneDrive/Desktop/EcoSys NEON Git")
```


#Tree Carbon Stocks-------------------------------------------
##############################################################

Environment
```{r}											
#install.packages("neonUtilities")						
library(neonUtilities)								
library(tidyverse)
library(lubridate)
```

#BLAN -- Tree Carbon Stocks
Downloading the data of interest from NEON

```{r}
 BLAN.trees <- loadByProduct(dpID="DP1.10098.001",
 site=c("BLAN"),
 startdate="2015-01", enddate="2024-12",
 package="basic")
 
 list2env(BLAN.trees, .GlobalEnv)
```

Write CSVs for four tables extracted
```{r}											
#readme_10098
write.csv(readme_10098, "readme_10098_BLAN.csv")

#vst_perplotperyear
write.csv(vst_perplotperyear, "vst_perplotperyear_BLAN.csv")

#vst_apparentindividual
write.csv(vst_apparentindividual, "vst_apparentindividual_BLAN.csv")

#variables_10098
write.csv(variables_10098, "variables_10098_BLAN.csv")
```

readme_10098 has a description of all the dataframes and variables_10098 defines all the variables in each dataframe. As you are working with NEON data in the future, you’ll want to familiarize yourself with these, but for the rest of this activity we will not need to refer to them.

Read in one of the CSVs we wrote above and revise to be compatible with the one that holds tree measurements
```{r}											
vst_perplotperyear_BLAN <- read.csv("vst_perplotperyear_BLAN.csv")
variables_10098_BLAN <- read.csv("variables_10098_BLAN.csv")

#making a new dataframe with selected variables
plot.info.vst.BLAN <- vst_perplotperyear_BLAN %>%
select(date, domainID, siteID, plotID, plotType, nlcdClass, totalSampledAreaTrees)

#making a new variable that containers only year
plot.info.vst.BLAN <- plot.info.vst.BLAN %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each PLOT
plot.info.vst.BLAN.rec <- plot.info.vst.BLAN %>%
group_by(plotID)%>%
filter(year==max(year))

#remove NAs from the totalSampledAreaTrees variable 
plot.info.vst.BLAN.rec <- plot.info.vst.BLAN.rec %>% filter(totalSampledAreaTrees!="NA")

```

Repeating the process above for individual dataframe
```{r}											
vst_apparentindividual_BLAN <- read.csv("vst_apparentindividual_BLAN.csv")

#making a new dataframe with selected variables
individuals.vst.BLAN <- vst_apparentindividual_BLAN %>%
select(date, siteID, plotID, individualID, growthForm, stemDiameter, plantStatus)

#making a new variable that containers only year
individuals.vst.BLAN <- individuals.vst.BLAN %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each INDIVIDUAL
most.recent.individuals.vst.BLAN <- individuals.vst.BLAN %>%
group_by(individualID) %>%
filter(year==max(year))

#making new dataframe where you filter the prior dataframe for only live trees
most.recent.live.plants.BLAN.vst <- most.recent.individuals.vst.BLAN %>% filter(str_detect(plantStatus, "Live")) #little unsure here

#making new dataframe for single bole trees, multi-bole trees, and small trees only
most.recent.live.trees1.BLAN <- most.recent.live.plants.BLAN.vst %>%
  filter(growthForm=="single bole tree" | growthForm=="multi-bole tree" | growthForm=="small tree")

#for dataframe above now we are excluding trees that have a recorded stem diameter of NA, we need to do this for analysis
most.recent.live.trees.BLAN <- subset(most.recent.live.trees1.BLAN, stemDiameter>0)

```

Joining dataframes for BLAN
```{r}

# str(most.recent.live.trees.BLAN)
# str(plot.info.vst.BLAN.rec)

tree.data.BLAN <- left_join(most.recent.live.trees.BLAN, plot.info.vst.BLAN.rec)
str(tree.data.BLAN) #595

tree.data.BLAN.complete<-subset(tree.data.BLAN, totalSampledAreaTrees>0)
# str(tree.data.BLAN.complete) #233

#We lost information for 362 trees when joining dataframes together.

```

```{r}

#need a new column in joined dataframe that represents biomass of each tree (in kg) called AGB
#mixed hardwood allometric equation from jenkins et al 2003
#Equation that I need to use is ABG (kg) = exp(B0 + B1 x ln(dbh))
tree.data.BLAN.complete$AGB <- exp(-2.4800 + (2.4835*log(tree.data.BLAN.complete$stemDiameter))) #this equation gives kg

#now need to create a new dataframe called AGB.trees.by.plot.BLAN
AGB.trees.by.plot.BLAN <- tree.data.BLAN.complete %>%
  group_by(siteID, plotID, totalSampledAreaTrees) %>% #grouping variables together
  summarise(plot.AGB = sum(AGB)) #summing by grouping variables 
# View(AGB.trees.by.plot.BLAN)

AGB.trees.by.plot.BLAN$AGB.kg.m2<-AGB.trees.by.plot.BLAN$plot.AGB/AGB.trees.by.plot.BLAN$totalSampledAreaTrees #making it as kg/m2

#Okay, need to get m2 to ha (10,000 m2 = 1 ha)
#then to get kg to mg (1,000 kg = 1 mg)

AGB.trees.by.plot.BLAN <- AGB.trees.by.plot.BLAN %>%
  mutate(AGB.kg.m2 = plot.AGB/totalSampledAreaTrees) %>%
  mutate(AGB.Mg.ha = (((AGB.kg.m2)/10))) %>%
  mutate(AGB.C.Mg.ha = AGB.Mg.ha*0.47)
 #View(AGB.trees.by.plot.BLAN)


```

```{r}
ggplot(data=AGB.trees.by.plot.BLAN, aes(x=plotID, y=AGB.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```


#SERC -- Tree Carbon Stocks
Downloading the data of interest from NEON
```{r}												
SERC.trees <- loadByProduct(dpID="DP1.10098.001", 	
site=c("SERC"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(SERC.trees, .GlobalEnv)
```	

Write CSVs for four tables extracted
```{r}											
#readme_10098
write.csv(readme_10098, "readme_10098_SERC.csv")

#vst_perplotperyear
write.csv(vst_perplotperyear, "vst_perplotperyear_SERC.csv")

#vst_apparentindividual
write.csv(vst_apparentindividual, "vst_apparentindividual_SERC.csv")

#variables_10098
write.csv(variables_10098, "variables_10098_SERC.csv")
```

readme_10098 has a description of all the dataframes and variables_10098 defines all the variables in each dataframe. As you are working with NEON data in the future, you’ll want to familiarize yourself with these, but for the rest of this activity we will not need to refer to them.

Read in one of the CSVs we wrote above and revise to be compatible with the one that holds tree measurements
```{r}											
vst_perplotperyear_SERC <- read.csv("vst_perplotperyear_SERC.csv")
variables_10098_SERC <- read.csv("variables_10098_SERC.csv")

#making a new dataframe with selected variables
plot.info.vst.SERC <- vst_perplotperyear_SERC %>%
select(date, domainID, siteID, plotID, plotType, nlcdClass, totalSampledAreaTrees)

#making a new variable that containers only year
plot.info.vst.SERC <- plot.info.vst.SERC %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each PLOT
plot.info.vst.SERC.rec <- plot.info.vst.SERC %>%
group_by(plotID)%>%
filter(year==max(year))

#remove NAs from the totalSampledAreaTrees variable 
plot.info.vst.SERC.rec <- plot.info.vst.SERC.rec %>% filter(totalSampledAreaTrees!="NA")

```

Repeating the process above for individual dataframe
```{r}											
vst_apparentindividual_SERC <- read.csv("vst_apparentindividual_SERC.csv")

#making a new dataframe with selected variables
individuals.vst.SERC <- vst_apparentindividual_SERC %>%
select(date, siteID, plotID, individualID, growthForm, stemDiameter, plantStatus)

#making a new variable that containers only year
individuals.vst.SERC <- individuals.vst.SERC %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each INDIVIDUAL
most.recent.individuals.vst.SERC <- individuals.vst.SERC %>%
group_by(individualID) %>%
filter(year==max(year))

#making new dataframe where you filter the prior dataframe for only live trees
most.recent.live.plants.SERC.vst <- most.recent.individuals.vst.SERC %>%
filter(str_detect(plantStatus, "Live")) #little unsure here

#making new dataframe for single bole trees, multi-bole trees, and small trees only
most.recent.live.trees1.SERC <- most.recent.live.plants.SERC.vst %>%
  filter(growthForm=="single bole tree" | growthForm=="multi-bole tree" | growthForm=="small tree")

#for dataframe above now we are excluding trees that have a recorded stem diameter of NA, we need to do this for analysis
most.recent.live.trees.SERC <- subset(most.recent.live.trees1.SERC, stemDiameter>0)

```

Joining dataframes for SERC
```{r}

# str(most.recent.live.trees.SERC)
# str(plot.info.vst.SERC.rec)

tree.data.SERC <- left_join(most.recent.live.trees.SERC, plot.info.vst.SERC.rec)
str(tree.data.SERC) #1753

tree.data.SERC.complete<-subset(tree.data.SERC, totalSampledAreaTrees>0)
str(tree.data.SERC.complete) #627

#We lost information for 1126 trees when joining dataframes together.

```

```{r}

#need a new column in joined dataframe that represents biomass of each tree (in kg) called AGB
#mixed hardwood allometric equation from jenkins et al 2003
#Equation that I need to use is ABG (kg) = exp(B0 + B1 x ln(dbh))
tree.data.SERC.complete$AGB <- exp(-2.4800 + (2.4835*log(tree.data.SERC.complete$stemDiameter))) #this equation gives kg

#now need to create a new dataframe called AGB.trees.by.plot.SERC
AGB.trees.by.plot.SERC <- tree.data.SERC.complete %>%
  group_by(siteID, plotID, totalSampledAreaTrees) %>% #grouping variables together
  summarise(plot.AGB = sum(AGB)) #summing by grouping variables 
View(AGB.trees.by.plot.SERC)

AGB.trees.by.plot.SERC$AGB.kg.m2<-AGB.trees.by.plot.SERC$plot.AGB/AGB.trees.by.plot.SERC$totalSampledAreaTrees #making it as kg/m2

#Okay, need to get m2 to ha (10,000 m2 = 1 ha)
#then to get kg to mg (1,000 kg = 1 mg)

AGB.trees.by.plot.SERC <- AGB.trees.by.plot.SERC %>%
  mutate(AGB.kg.m2 = plot.AGB/totalSampledAreaTrees) %>%
  mutate(AGB.Mg.ha = (((AGB.kg.m2)*10))) %>%
  mutate(AGB.C.Mg.ha = AGB.Mg.ha*0.47)
# View(AGB.trees.by.plot.SERC)

```

```{r}
ggplot(data=AGB.trees.by.plot.SERC, aes(x=plotID, y=AGB.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#SCBI -- Tree Carbon Stocks
Downloading the data of interest from NEON
```{r}												
SCBI.trees <- loadByProduct(dpID="DP1.10098.001", 	
site=c("SCBI"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(SCBI.trees, .GlobalEnv)
```	

Write CSVs for four tables extracted
```{r}											
#readme_10098
write.csv(readme_10098, "readme_10098_SCBI.csv")

#vst_perplotperyear
write.csv(vst_perplotperyear, "vst_perplotperyear_SCBI.csv")

#vst_apparentindividual
 write.csv(vst_apparentindividual, "vst_apparentindividual_SCBI.csv")

#variables_10098
write.csv(variables_10098, "variables_10098_SCBI.csv")
```

readme_10098 has a description of all the dataframes and variables_10098 defines all the variables in each dataframe. As you are working with NEON data in the future, you’ll want to familiarize yourself with these, but for the rest of this activity we will not need to refer to them.

Read in one of the CSVs we wrote above and revise to be compatible with the one that holds tree measurements
```{r}											
vst_perplotperyear_SCBI <- read.csv("vst_perplotperyear_SCBI.csv")
variables_10098_SCBI <- read.csv("variables_10098_SCBI.csv")

#making a new dataframe with selected variables
plot.info.vst.SCBI <- vst_perplotperyear_SCBI %>%
select(date, domainID, siteID, plotID, plotType, nlcdClass, totalSampledAreaTrees)

#making a new variable that containers only year
plot.info.vst.SCBI <- plot.info.vst.SCBI %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each PLOT
plot.info.vst.SCBI.rec <- plot.info.vst.SCBI %>%
group_by(plotID)%>%
filter(year==max(year))

#remove NAs from the totalSampledAreaTrees variable 
plot.info.vst.SCBI.rec <- plot.info.vst.SCBI.rec %>% filter(totalSampledAreaTrees!="NA")

```

Repeating the process above for individual dataframe
```{r}											
vst_apparentindividual_SCBI <- read.csv("vst_apparentindividual_SCBI.csv")

#making a new dataframe with selected variables
individuals.vst.SCBI <- vst_apparentindividual_SCBI %>%
select(date, siteID, plotID, individualID, growthForm, stemDiameter, plantStatus)

#making a new variable that containers only year
individuals.vst.SCBI <- individuals.vst.SCBI %>%
mutate(year= year(date))

#making new dataframe where you filter only the most recent sampling year for each INDIVIDUAL
most.recent.individuals.vst.SCBI <- individuals.vst.SCBI %>%
group_by(individualID) %>%
filter(year==max(year))

#making new dataframe where you filter the prior dataframe for only live trees
most.recent.live.plants.SCBI.vst <- most.recent.individuals.vst.SCBI %>%
filter(str_detect(plantStatus, "Live")) #little unsure here

#making new dataframe for single bole trees, multi-bole trees, and small trees only
most.recent.live.trees1.SCBI <- most.recent.live.plants.SCBI.vst %>%
  filter(growthForm=="single bole tree" | growthForm=="multi-bole tree" | growthForm=="small tree")

#for dataframe above now we are excluding trees that have a recorded stem diameter of NA, we need to do this for analysis
most.recent.live.trees.SCBI <- subset(most.recent.live.trees1.SCBI, stemDiameter>0)

```

Joining dataframes for SCBI
```{r}

# str(most.recent.live.trees.SCBI)
# str(plot.info.vst.SCBI.rec)

tree.data.SCBI <- left_join(most.recent.live.trees.SCBI, plot.info.vst.SCBI.rec)
str(tree.data.SCBI) #962

tree.data.SCBI.complete<-subset(tree.data.SCBI, totalSampledAreaTrees>0)
str(tree.data.SCBI.complete) #399

#We lost information for 563 trees when joining dataframes together.

```

```{r}

#need a new column in joined dataframe that represents biomass of each tree (in kg) called AGB
#mixed hardwood allometric equation from jenkins et al 2003
#Equation that I need to use is ABG (kg) = exp(B0 + B1 x ln(dbh))
tree.data.SCBI.complete$AGB <- exp(-2.4800 + (2.4835*log(tree.data.SCBI.complete$stemDiameter))) #this equation gives kg

#now need to create a new dataframe called AGB.trees.by.plot.SCBI
AGB.trees.by.plot.SCBI <- tree.data.SCBI.complete %>%
  group_by(siteID, plotID, totalSampledAreaTrees) %>% #grouping variables together
  summarise(plot.AGB = sum(AGB)) #summing by grouping variables 
# View(AGB.trees.by.plot.SCBI)

AGB.trees.by.plot.SCBI$AGB.kg.m2<-AGB.trees.by.plot.SCBI$plot.AGB/AGB.trees.by.plot.SCBI$totalSampledAreaTrees #making it as kg/m2

#Okay, need to get m2 to ha (10,000 m2 = 1 ha)
#then to get kg to mg (1,000 kg = 1 mg)

AGB.trees.by.plot.SCBI <- AGB.trees.by.plot.SCBI %>%
  mutate(AGB.kg.m2 = plot.AGB/totalSampledAreaTrees) %>%
  mutate(AGB.Mg.ha = (((AGB.kg.m2)*10))) %>%
  mutate(AGB.C.Mg.ha = AGB.Mg.ha*0.47)
# View(AGB.trees.by.plot.SCBI)

```

```{r}
ggplot(data=AGB.trees.by.plot.SCBI, aes(x=plotID, y=AGB.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```



#Joining dataframes across D02 (Mid-Atlantic) for Tree Carbon Stocks
```{r}

str(AGB.trees.by.plot.BLAN) #10
str(AGB.trees.by.plot.SERC) #30
str(AGB.trees.by.plot.SCBI) #26

AGB.trees.by.plot.D02 <- full_join(AGB.trees.by.plot.BLAN, AGB.trees.by.plot.SERC)
AGB.trees.by.plot.D02 <- full_join(AGB.trees.by.plot.D02, AGB.trees.by.plot.SCBI)
str(AGB.trees.by.plot.D02)  #66
```

#Visualizing Tree Carbon Stocks by site of D02
```{r}

AGB.trees.by.plot.D02.summarise <- AGB.trees.by.plot.D02 %>%
 group_by(siteID) %>%
 summarise(AGB.C.Mg.ha_siteAverage = mean(AGB.C.Mg.ha),
           AGB.C.Mg.ha_siteSD = sd(AGB.C.Mg.ha))
 
ggplot(data=AGB.trees.by.plot.D02, aes(x=siteID, y=AGB.C.Mg.ha, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = AGB.trees.by.plot.D02.summarise, aes(x = siteID, y = AGB.C.Mg.ha_siteAverage, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = AGB.trees.by.plot.D02.summarise, aes(x= siteID, y = AGB.C.Mg.ha_siteAverage, ymin = (AGB.C.Mg.ha_siteAverage-AGB.C.Mg.ha_siteSD), ymax = (AGB.C.Mg.ha_siteAverage+AGB.C.Mg.ha_siteSD), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Tree Carbon Stocks by Site of D02") +
        ylab("Carbon Stock (Mg/ha)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none") 

```


#Root Carbon Stocks-------------------------------------------
##############################################################

#BLAN -- Root Carbon Stocks
Create a chunk where you extract root biomass using NEON utilies package, get into environment, and write csvs into the environment
```{r}

BLAN.roots <- loadByProduct(dpID="DP1.10067.001", 	
site=c("BLAN"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(BLAN.roots, .GlobalEnv)

#readme_10067
write.csv(readme_10067, "readme_10067_BLAN.csv")

#bbc_percore
write.csv(bbc_percore, "bbc_percore_BLAN.csv")

#bbc_rootmass
write.csv(bbc_rootmass, "bbc_rootmass_BLAN.csv")

#variables_10067
write.csv(variables_10067, "variables_10067_BLAN.csv")

```

Create chunk selecting variables to create new dfs
```{r}

bbc_rootmass_BLAN <- read.csv("bbc_rootmass_BLAN.csv")
bbc_percore_BLAN <- read.csv("bbc_percore_BLAN.csv")

root.mass.BLAN <- bbc_rootmass_BLAN %>%
select(domainID, plotID, sampleID, subsampleID, collectDate, sizeCategory, rootStatus, dryMass)

root.core.BLAN <- bbc_percore_BLAN %>%
 select(domainID, plotID, subplotID, sampleID, clipID, coreID, collectDate, rootSampleArea, rootSampleDepth, coreDiameter)


root.data.BLAN <- left_join(root.mass.BLAN, root.core.BLAN)

root.data.BLAN$year <- substr(root.data.BLAN$collectDate, start=1, stop=4) 

```

Root calculations and conversions
```{r}

roots.by.sample.BLAN <- root.data.BLAN %>% 
  group_by(plotID, sampleID, rootSampleArea, rootSampleDepth) %>%
  summarise(total.mass = sum(dryMass))

roots.by.sample.BLAN.final <- roots.by.sample.BLAN %>%
  mutate(root.gC.sample = (total.mass * 0.47)) %>%
  mutate(root.gC.m2 = (root.gC.sample/rootSampleArea)) %>%
  mutate(root.MgC.ha = (root.gC.m2 * 0.01))
 
roots.by.plot.BLAN <- roots.by.sample.BLAN.final %>% 
  group_by(plotID) %>%
  summarise(mean.root.MgC.ha = mean(root.MgC.ha))

```

Visualize the data
```{r}

ggplot(data=roots.by.plot.BLAN, aes(x=plotID, y=mean.root.MgC.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#SERC -- Root Carbon Stocks
Create a chunk where you extract root biomass using NEON utilies package, get into environment, and write csvs into the environment
```{r}

SERC.roots <- loadByProduct(dpID="DP1.10067.001",
site=c("SERC"),
startdate="2012-01", enddate="2024-12",
package="basic")

list2env(SERC.roots, .GlobalEnv)

#readme_10067
write.csv(readme_10067, "readme_10067_SERC.csv")

#bbc_percore
write.csv(bbc_percore, "bbc_percore_SERC.csv")

#bbc_rootmass
write.csv(bbc_rootmass, "bbc_rootmass_SERC.csv")

#variables_10067
write.csv(variables_10067, "variables_10067_SERC.csv")

```

Create chunk selecting variables to create new dfs
```{r}

bbc_rootmass_SERC <- read.csv("bbc_rootmass_SERC.csv")
bbc_percore_SERC <- read.csv("bbc_percore_SERC.csv")

root.mass.SERC <- bbc_rootmass_SERC %>%
select(domainID, plotID, sampleID, subsampleID, collectDate, sizeCategory, rootStatus, dryMass)

root.core.SERC <- bbc_percore_SERC %>%
 select(domainID, plotID, subplotID, sampleID, clipID, coreID, collectDate, rootSampleArea, rootSampleDepth, coreDiameter)


root.data.SERC <- left_join(root.mass.SERC, root.core.SERC)

root.data.SERC$year <- substr(root.data.SERC$collectDate, start=1, stop=4) 

```

Root calculations and conversions
```{r}

roots.by.sample.SERC <- root.data.SERC %>% 
  group_by(plotID, sampleID, rootSampleArea, rootSampleDepth) %>%
  summarise(total.mass = sum(dryMass))

roots.by.sample.SERC.final <- roots.by.sample.SERC %>%
  mutate(root.gC.sample = (total.mass * 0.47)) %>%
  mutate(root.gC.m2 = (root.gC.sample/rootSampleArea)) %>%
  mutate(root.MgC.ha = (root.gC.m2 * 0.01))
 
roots.by.plot.SERC <- roots.by.sample.SERC.final %>% 
  group_by(plotID) %>%
  summarise(mean.root.MgC.ha = mean(root.MgC.ha))

```

Visualize the data
```{r}

ggplot(data=roots.by.plot.SERC, aes(x=plotID, y=mean.root.MgC.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#SCBI -- Root Carbon Stocks
Create a chunk where you extract root biomass using NEON utilies package, get into environment, and write csvs into the environment
```{r}

SCBI.roots <- loadByProduct(dpID="DP1.10067.001", 	
site=c("SCBI"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

 list2env(SCBI.roots, .GlobalEnv)

#readme_10067
write.csv(readme_10067, "readme_10067_SCBI.csv")

#bbc_percore
write.csv(bbc_percore, "bbc_percore_SCBI.csv")

#bbc_rootmass
write.csv(bbc_rootmass, "bbc_rootmass_SCBI.csv")

#variables_10067
write.csv(variables_10067, "variables_10067_SCBI.csv")

```

Create chunk selecting variables to create new dfs
```{r}

bbc_rootmass_SCBI <- read.csv("bbc_rootmass_SCBI.csv")
bbc_percore_SCBI <- read.csv("bbc_percore_SCBI.csv")

root.mass.SCBI <- bbc_rootmass_SCBI %>%
select(domainID, plotID, sampleID, subsampleID, collectDate, sizeCategory, rootStatus, dryMass)

root.core.SCBI <- bbc_percore_SCBI %>%
 select(domainID, plotID, subplotID, sampleID, clipID, coreID, collectDate, rootSampleArea, rootSampleDepth, coreDiameter)


root.data.SCBI <- left_join(root.mass.SCBI, root.core.SCBI)

root.data.SCBI$year <- substr(root.data.SCBI$collectDate, start=1, stop=4) 

```

Root calculations and conversions
```{r}

roots.by.sample.SCBI <- root.data.SCBI %>% 
  group_by(plotID, sampleID, rootSampleArea, rootSampleDepth) %>%
  summarise(total.mass = sum(dryMass))

roots.by.sample.SCBI.final <- roots.by.sample.SCBI %>%
  mutate(root.gC.sample = (total.mass * 0.47)) %>%
  mutate(root.gC.m2 = (root.gC.sample/rootSampleArea)) %>%
  mutate(root.MgC.ha = (root.gC.m2 * 0.01))
 
roots.by.plot.SCBI <- roots.by.sample.SCBI.final %>% 
  group_by(plotID) %>%
  summarise(mean.root.MgC.ha = mean(root.MgC.ha))

```

Visualize the data
```{r}

ggplot(data=roots.by.plot.SCBI, aes(x=plotID, y=mean.root.MgC.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```

#Joining dataframes across D02 (Southeastern longleaf pine forests) for Root Carbon Stocks
```{r}

str(roots.by.plot.BLAN) #30
str(roots.by.plot.SERC) #20
str(roots.by.plot.SCBI) #20

roots.by.plot.D02 <- full_join(roots.by.plot.BLAN, roots.by.plot.SERC)
roots.by.plot.D02 <- full_join(roots.by.plot.D02, roots.by.plot.SCBI)
str(roots.by.plot.D02)  #70
```

#Visualizing Root Carbon Stocks by site of D02
```{r}

roots.by.plot.D02$siteID <- str_sub(roots.by.plot.D02$plotID, 1, 4)

roots.by.plot.D02.summarise <- roots.by.plot.D02 %>%
 group_by(siteID) %>%
 summarise(mean.root.MgC.ha_siteAverage = mean(mean.root.MgC.ha),
           mean.root.MgC.ha_siteSD = sd(mean.root.MgC.ha)) 
 
ggplot(data=roots.by.plot.D02, aes(x=siteID, y=mean.root.MgC.ha, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = roots.by.plot.D02.summarise, aes(x = siteID, y = mean.root.MgC.ha_siteAverage, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = roots.by.plot.D02.summarise, aes(x= siteID, y = mean.root.MgC.ha_siteAverage, ymin = (mean.root.MgC.ha_siteAverage-mean.root.MgC.ha_siteSD), ymax = (mean.root.MgC.ha_siteAverage+mean.root.MgC.ha_siteSD), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Root Carbon Stocks by Site of D02") +
        ylab("Carbon Stock (Mg/ha)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none") 

```




```{r}
R.version
```
#Soil Carbon Stock--------------------------------------------
##############################################################

#BLAN -- Soil Carbon Stocks

Downloading the data of interest from NEON, lising, and csvs
```{r}												
BLAN.soil <- loadByProduct(dpID="DP1.10047.001", 	
site=c("BLAN"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(BLAN.soil, .GlobalEnv)

#spc_biogeochem
write.csv(spc_biogeochem, "spc_biogeochem_BLAN.csv")

#spc_bulkdensity
write.csv(spc_bulkdensity, "spc_bulkdensity_BLAN.csv")
```

To quantify soil carbon stocks, we need bulk density and % carbon values for each soil increment. Inconveniently, NEON does not put these variables into the same data set, so we need to perform some carpentry. All the steps below are similar to what you performed with trees and roots.

1a-3d
```{r}											
spc_biogeochem_BLAN <- read.csv("spc_biogeochem_BLAN.csv")
spc_bulkdensity_BLAN <- read.csv("spc_bulkdensity_BLAN.csv")

#making a new dataframe with selected variables
biogeochem.info.BLAN <- spc_biogeochem_BLAN %>%
select(domainID, siteID, plotID, collectDate, horizonID, horizonName, biogeoTopDepth, biogeoBottomDepth, carbonTot)

#making a new dataframe with selected variables
bulkdensity.info.BLAN <- spc_bulkdensity_BLAN %>%
select(domainID, siteID, plotID, collectDate, horizonID, horizonName, bulkDensThirdBar, bulkDensFieldMoist)

soil.data.BLAN <- left_join(biogeochem.info.BLAN, bulkdensity.info.BLAN)
str(soil.data.BLAN) #60 observations, 11 variables

soil.data.BLAN <- soil.data.BLAN %>%
mutate(depth_increment = biogeoBottomDepth-biogeoTopDepth)
str(soil.data.BLAN) #60 observations, 12 variables
```

e)	GAP FILL View your data frame again. Notice how many NAs there are for the bulk density variable! This is a common problem—bulk density measures are rare. We will deal with this issue by ‘gap filling’.


```{r}
sum(is.na(soil.data.BLAN$bulkDensThirdBar)) #11 observations for NA
sum(is.na(soil.data.BLAN$bulkDensFieldMoist)) #60 observations for NA

#(2)	First calculate the mean bulk density (using bulkDensFieldMoist, since it’s more abundant) for each plot and then for each site. Perform the following steps:

#(a) Use mutate to create a new variable in your existing dataframe of the mean bulk density of each plot. 
soil.data.BLAN <- soil.data.BLAN %>% 
 group_by(plotID) %>% 
 mutate(BD.plot.mean = mean(bulkDensThirdBar, na.rm=TRUE))
str(soil.data.BLAN)  #60 obs. 20 var.

#(b)	Use mutate to calculate the mean bulk density of each site. Use similar steps as above, but group by siteID instead of plotID.
soil.data.BLAN <- soil.data.BLAN %>% 
 group_by(siteID) %>% 
 mutate(BD.site.mean = mean(bulkDensThirdBar, na.rm=TRUE))
str(soil.data.BLAN) #60 obs 20 var

#(3)	Gap fill using if_else to fill in values from bulkDensFieldMoist where bulkDensThirdBar is absent. Use this logic: if the NEON provided bulk density from third bar method is NA, apply the value from the FieldMoist method to the new variable ‘bulkdensity’, else, apply the ThirdBar value to the new variable ‘bulkdensity’.

soil.data.BLAN <- soil.data.BLAN %>% 
group_by(plotID) %>% 
mutate(bulkdensity = if_else(is.na(bulkDensFieldMoist), bulkDensThirdBar , bulkDensFieldMoist))

sum(is.na(soil.data.BLAN$bulkdensity)) #how many NAs? 11 


#(4)	If you still have NAs use the following steps to backfill using this logic: if the NEON provided bulk density value is NA, apply the plot-level mean to the variable ‘bulkdensity’, else, apply the NEON provide value to the variable ‘bulkdensity’.

#####NOTE#####Was having issues getting rid of NAs after back and gap filling. Decided to go back and have the mean depend on bulkdensthirdbar since there bulkdensfieldmoist wasnt working. Then decided to swap bulkdensthirdbar with bulkdensity in the following code.

soil.data.BLAN <- soil.data.BLAN %>%
group_by(plotID) %>%
mutate(bulkdensity = if_else(is.na(bulkdensity), bulkDensThirdBar, bulkdensity))

sum(is.na(soil.data.BLAN$bulkdensity)) #how many NAs? 11

#(5) Gap fill again with site-level mean. Use the if_else approach with this logic: if the new bulkdensity variable is NA, apply the site-level mean to the variable ‘bulkdensity,’ else apply the value already given to ‘bulkdensity’.

soil.data.BLAN <- soil.data.BLAN %>%
group_by(plotID) %>%
mutate(bulkdensity = if_else(is.na(bulkdensity), BD.site.mean, bulkdensity))

sum(is.na(soil.data.BLAN$bulkdensity)) #how many NAs? 0

View(soil.data.BLAN)

```

f)	Calculate the soil carbon stock
```{r}
#(1)	Given bulk density and the depth increment of each horizon in your dataframe, create a new column in your dataframe that is the total amount of soil in grams for a m^2 area

#(a-d)	To do this, go back to your HSB code and notice how we calculated the sample volume for each depth increment. We used a m2 area, so use 100 cm x 100 cm  = 10,000 cm2 for the sample area in this case too.

soil.data.BLAN <- soil.data.BLAN %>% 
  mutate(sample.area.cm2 = 10000) %>%
  mutate(sample.volume = sample.area.cm2*depth_increment) %>%  #created an arbitrary variable of 10,000 cm2 that represents 1 m2 and then multiplying by the depth increment we calculated above to try to get volume, which will be now in cm3
  mutate(g_soil = sample.volume*bulkdensity) %>% #this is in g/m2
  mutate(kg_m2_soil = g_soil/1000) #should now be in kg/m2

#what are the units of carbonTots? According to variables_10047, units are gramsPerKilogram, g/kg

#(e-f) Calculate the total carbon by each depth increment... not using increment directly, but essentially using the kilograms per m2 of soil calculated above, since that represents the kilograms of soil at a given depth_increment within a m2 area.
soil.data.BLAN <- soil.data.BLAN %>% 
  mutate(carbon_stock_forDepth = carbonTot*kg_m2_soil) #units in g/m2

#(g-h) Create a new dataframe that sums the carbon stock of all the depth increments
soil.data.sum.BLAN <- soil.data.BLAN %>% 
  group_by(plotID, siteID) %>%
  summarise(total.carbon = sum(carbon_stock_forDepth))

#(h) Convert to Mg C per ha
soil.data.sum.BLAN <- soil.data.sum.BLAN %>% 
  mutate(soil.C.Mg.ha = (total.carbon*0.01))
  
  
#(i) Visualize the data
ggplot(data=soil.data.sum.BLAN, aes(x=plotID, y=soil.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


```


#SERC -- Soil Carbon Stocks

Downloading the data of interest from NEON, lising, and csvs
```{r}												
SERC.soil <- loadByProduct(dpID="DP1.10047.001", 	
site=c("SERC"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(SERC.soil, .GlobalEnv)

#spc_biogeochem
write.csv(spc_biogeochem, "spc_biogeochem_SERC.csv")

#spc_bulkdensity
write.csv(spc_bulkdensity, "spc_bulkdensity_SERC.csv")
```

To quantify soil carbon stocks, we need bulk density and % carbon values for each soil increment. Inconveniently, NEON does not put these variables into the same data set, so we need to perform some carpentry. All the steps below are similar to what you performed with trees and roots.

1a-3d
```{r}											
spc_biogeochem_SERC <- read.csv("spc_biogeochem_SERC.csv")
spc_bulkdensity_SERC <- read.csv("spc_bulkdensity_SERC.csv")

#making a new dataframe with selected variables
biogeochem.info.SERC <- spc_biogeochem_SERC %>%
select(domainID, siteID, plotID, collectDate, horizonID, horizonName, biogeoTopDepth, biogeoBottomDepth, carbonTot)

#making a new dataframe with selected variables
bulkdensity.info.SERC <- spc_bulkdensity_SERC %>%
select(domainID, siteID, plotID, collectDate, horizonID, horizonName, bulkDensThirdBar, bulkDensFieldMoist)

soil.data.SERC <- left_join(biogeochem.info.SERC, bulkdensity.info.SERC)
str(soil.data.SERC) #92 observations, 11 variables

soil.data.SERC <- soil.data.SERC %>%
mutate(depth_increment = biogeoBottomDepth-biogeoTopDepth)
str(soil.data.SERC) #92 observations, 12 variables
```

e)	GAP FILL View your data frame again. Notice how many NAs there are for the bulk density variable! This is a common problem—bulk density measures are rare. We will deal with this issue by ‘gap filling’.


```{r}
sum(is.na(soil.data.SERC$bulkDensThirdBar)) #3 observations for NA
sum(is.na(soil.data.SERC$bulkDensFieldMoist)) #89 observations for NA

#(2)	First calculate the mean bulk density (using bulkDensFieldMoist, since it’s more abundant) for each plot and then for each site. Perform the following steps:

#(a) Use mutate to create a new variable in your existing dataframe of the mean bulk density of each plot. 
soil.data.SERC <- soil.data.SERC %>% 
 group_by(plotID) %>% 
 mutate(BD.plot.mean = mean(bulkDensFieldMoist, na.rm=TRUE))
str(soil.data.SERC) #92 obs. 13 var

#(b)	Use mutate to calculate the mean bulk density of each site. Use similar steps as above, but group by siteID instead of plotID.
soil.data.SERC <- soil.data.SERC %>% 
 group_by(siteID) %>% 
 mutate(BD.site.mean = mean(bulkDensFieldMoist, na.rm=TRUE))


#(3)	Gap fill using if_else to fill in values from bulkDensFieldMoist where bulkDensThirdBar is absent. Use this logic: if the NEON provided bulk density from third bar method is NA, apply the value from the FieldMoist method to the new variable ‘bulkdensity’, else, apply the ThirdBar value to the new variable ‘bulkdensity’.

soil.data.SERC <- soil.data.SERC %>% 
group_by(plotID) %>% 
mutate(bulkdensity = if_else(is.na(bulkDensThirdBar), bulkDensFieldMoist , bulkDensThirdBar))

sum(is.na(soil.data.SERC$bulkdensity)) #how many NAs? 0

#(4)	If you still have NAs use the following steps to backfill using this logic: if the NEON provided bulk density value is NA, apply the plot-level mean to the variable ‘bulkdensity’, else, apply the NEON provide value to the variable ‘bulkdensity’.

soil.data.SERC <- soil.data.SERC %>%
group_by(plotID) %>%
mutate(bulkdensity = if_else(is.na(bulkdensity), BD.plot.mean, bulkdensity))

sum(is.na(soil.data.SERC$bulkdensity)) #how many NAs? #0

#(5) Gap fill again with site-level mean. Use the if_else approach with this logic: if the new bulkdensity variable is NA, apply the site-level mean to the variable ‘bulkdensity,’ else apply the value already given to ‘bulkdensity’.

soil.data.SERC <- soil.data.SERC %>%
group_by(plotID) %>%
mutate(bulkdensity = if_else(is.na(bulkdensity), BD.site.mean, bulkdensity))

sum(is.na(soil.data.SERC$bulkdensity)) #how many NAs? 0

# View(soil.data.SERC)

```

f)	Calculate the soil carbon stock
```{r}
#(1)	Given bulk density and the depth increment of each horizon in your dataframe, create a new column in your dataframe that is the total amount of soil in grams for a m^2 area

#(a-d)	To do this, go back to your HSB code and notice how we calculated the sample volume for each depth increment. We used a m2 area, so use 100 cm x 100 cm  = 10,000 cm2 for the sample area in this case too.

soil.data.SERC <- soil.data.SERC %>% 
  mutate(sample.area.cm2 = 10000) %>%
  mutate(sample.volume = sample.area.cm2*depth_increment) %>%  #created an arbitrary variable of 10,000 cm2 that represents 1 m2 and then multiplying by the depth increment we calculated above to try to get volume, which will be now in cm3
  mutate(g_soil = sample.volume*bulkdensity) %>% #this is in g/m2
  mutate(kg_m2_soil = g_soil/1000) #should now be in kg/m2

#what are the units of carbonTots? According to variables_10047, units are gramsPerKilogram, g/kg

#(e-f) Calculate the total carbon by each depth increment... not using increment directly, but essentially using the kilograms per m2 of soil calculated above, since that represents the kilograms of soil at a given depth_increment within a m2 area.
soil.data.SERC <- soil.data.SERC %>% 
  mutate(carbon_stock_forDepth = carbonTot*kg_m2_soil) #units in g/m2

#(g-h) Create a new dataframe that sums the carbon stock of all the depth increments
soil.data.sum.SERC <- soil.data.SERC %>% 
  group_by(plotID, siteID) %>%
  summarise(total.carbon = sum(carbon_stock_forDepth))

#(h) Convert to Mg C per ha
soil.data.sum.SERC <- soil.data.sum.SERC %>% 
  mutate(soil.C.Mg.ha = (total.carbon*0.01))
  
  
#(i) Visualize the data
ggplot(data=soil.data.sum.SERC, aes(x=plotID, y=soil.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


```

#SCBI -- Soil Carbon Stocks

Downloading the data of interest from NEON, lising, and csvs
```{r}												
SCBI.soil <- loadByProduct(dpID="DP1.10047.001", 	
site=c("SCBI"), 		
startdate="2012-01", enddate="2024-12", 				
package="basic")

list2env(SCBI.soil, .GlobalEnv)

#spc_biogeochem
write.csv(spc_biogeochem, "spc_biogeochem_SCBI.csv")

#spc_bulkdensity
write.csv(spc_bulkdensity, "spc_bulkdensity_SCBI.csv")
```

To quantify soil carbon stocks, we need bulk density and % carbon values for each soil increment. Inconveniently, NEON does not put these variables into the same data set, so we need to perform some carpentry. All the steps below are similar to what you performed with trees and roots.

1a-3d
```{r}											
spc_biogeochem_SCBI <- read.csv("spc_biogeochem_SCBI.csv")
spc_bulkdensity_SCBI <- read.csv("spc_bulkdensity_SCBI.csv")

#making a new dataframe with selected variables
biogeochem.info.SCBI <- spc_biogeochem_SCBI %>%
select(domainID, siteID, plotID, collectDate, horizonID, horizonName, biogeoTopDepth, biogeoBottomDepth, carbonTot)

#making a new dataframe with selected variables
bulkdensity.info.SCBI <- spc_bulkdensity_SCBI %>%
select(domainID, siteID, plotID, collectDate, horizonID, horizonName, bulkDensThirdBar, bulkDensFieldMoist)

soil.data.SCBI <- left_join(biogeochem.info.SCBI, bulkdensity.info.SCBI)
str(soil.data.SCBI) #89 observations, 11 variables

soil.data.SCBI <- soil.data.SCBI %>%
mutate(depth_increment = biogeoBottomDepth-biogeoTopDepth)
str(soil.data.SCBI) #89 observations, 12 variables
```

e)	GAP FILL View your data frame again. Notice how many NAs there are for the bulk density variable! This is a common problem—bulk density measures are rare. We will deal with this issue by ‘gap filling’.


```{r}
sum(is.na(soil.data.SCBI$bulkDensThirdBar)) #21 observations for NA
sum(is.na(soil.data.SCBI$bulkDensFieldMoist)) #84 observations for NA

#(2)	First calculate the mean bulk density (using bulkDensFieldMoist, since it’s more abundant) for each plot and then for each site. Perform the following steps:

#(a) Use mutate to create a new variable in your existing dataframe of the mean bulk density of each plot. 
soil.data.SCBI <- soil.data.SCBI %>% 
 group_by(plotID) %>% 
 mutate(BD.plot.mean = mean(bulkDensFieldMoist, na.rm=TRUE))
str(soil.data.SCBI)  #89 obs 13 var

#(b)	Use mutate to calculate the mean bulk density of each site. Use similar steps as above, but group by siteID instead of plotID.
soil.data.SCBI <- soil.data.SCBI %>% 
 group_by(siteID) %>% 
 mutate(BD.site.mean = mean(bulkDensFieldMoist, na.rm=TRUE))


#(3)	Gap fill using if_else to fill in values from bulkDensFieldMoist where bulkDensThirdBar is absent. Use this logic: if the NEON provided bulk density from third bar method is NA, apply the value from the FieldMoist method to the new variable ‘bulkdensity’, else, apply the ThirdBar value to the new variable ‘bulkdensity’.

soil.data.SCBI <- soil.data.SCBI %>% 
group_by(plotID) %>% 
mutate(bulkdensity = if_else(is.na(bulkDensThirdBar), bulkDensFieldMoist , bulkDensThirdBar))

sum(is.na(soil.data.SCBI$bulkdensity)) #how many NAs? 16

#(4)	If you still have NAs use the following steps to backfill using this logic: if the NEON provided bulk density value is NA, apply the plot-level mean to the variable ‘bulkdensity’, else, apply the NEON provide value to the variable ‘bulkdensity’.

soil.data.SCBI <- soil.data.SCBI %>%
group_by(plotID) %>%
mutate(bulkdensity = if_else(is.na(bulkdensity), BD.plot.mean, bulkdensity))

sum(is.na(soil.data.SCBI$bulkdensity)) #how many NAs? 14

#(5) Gap fill again with site-level mean. Use the if_else approach with this logic: if the new bulkdensity variable is NA, apply the site-level mean to the variable ‘bulkdensity,’ else apply the value already given to ‘bulkdensity’.

soil.data.SCBI <- soil.data.SCBI %>%
group_by(plotID) %>%
mutate(bulkdensity = if_else(is.na(bulkdensity), BD.site.mean, bulkdensity))

sum(is.na(soil.data.SCBI$bulkdensity)) #how many NAs? 0

# View(soil.data.SCBI)

```

f)	Calculate the soil carbon stock
```{r}
#(1)	Given bulk density and the depth increment of each horizon in your dataframe, create a new column in your dataframe that is the total amount of soil in grams for a m^2 area

#(a-d)	To do this, go back to your HSB code and notice how we calculated the sample volume for each depth increment. We used a m2 area, so use 100 cm x 100 cm  = 10,000 cm2 for the sample area in this case too.

soil.data.SCBI <- soil.data.SCBI %>% 
  mutate(sample.area.cm2 = 10000) %>%
  mutate(sample.volume = sample.area.cm2*depth_increment) %>%  #created an arbitrary variable of 10,000 cm2 that represents 1 m2 and then multiplying by the depth increment we calculated above to try to get volume, which will be now in cm3
  mutate(g_soil = sample.volume*bulkdensity) %>% #this is in g/m2
  mutate(kg_m2_soil = g_soil/1000) #should now be in kg/m2

#what are the units of carbonTots? According to variables_10047, units are gramsPerKilogram, g/kg

#(e-f) Calculate the total carbon by each depth increment... not using increment directly, but essentially using the kilograms per m2 of soil calculated above, since that represents the kilograms of soil at a given depth_increment within a m2 area.
soil.data.SCBI <- soil.data.SCBI %>% 
  mutate(carbon_stock_forDepth = carbonTot*kg_m2_soil) #units in g/m2

#(g-h) Create a new dataframe that sums the carbon stock of all the depth increments
soil.data.sum.SCBI <- soil.data.SCBI %>% 
  group_by(plotID, siteID) %>%
  summarise(total.carbon = sum(carbon_stock_forDepth))

#(h) Convert to Mg C per ha
soil.data.sum.SCBI <- soil.data.sum.SCBI %>% 
  mutate(soil.C.Mg.ha = (total.carbon*0.01))
  
  
#(i) Visualize the data
ggplot(data=soil.data.sum.SCBI, aes(x=plotID, y=soil.C.Mg.ha)) + 
geom_jitter()+
theme_classic()+
theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


```

#Joining dataframes across D02 (Southeastern longleaf pine forests) for Soil Carbon Stocks
```{r}

str(soil.data.sum.BLAN) #15 obs 4 var
str(soil.data.sum.SERC) #20 obs 4 var
str(soil.data.sum.SCBI) #18 obs 4 var

soil.by.plot.D02 <- full_join(soil.data.sum.BLAN, soil.data.sum.SERC)
soil.by.plot.D02 <- full_join(soil.by.plot.D02, soil.data.sum.SCBI)
str(roots.by.plot.D02)  #53 obs 4 var
```

#Visualizing Soil Carbon Stocks by site of D02
```{r}

soil.by.plot.D02.summarise <- soil.by.plot.D02 %>%
 group_by(siteID) %>%
 summarise(soil.C.Mg.ha_siteAverage = mean(soil.C.Mg.ha),
           soil.C.Mg.ha_siteSD = sd(soil.C.Mg.ha)) 
 
ggplot(data=soil.by.plot.D02, aes(x=siteID, y=soil.C.Mg.ha, group = siteID, color=siteID)) + 
  geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
  geom_point(data = soil.by.plot.D02.summarise, aes(x = siteID, y = soil.C.Mg.ha_siteAverage, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
  geom_errorbar(data = soil.by.plot.D02.summarise, aes(x= siteID, y = soil.C.Mg.ha_siteAverage, ymin = (soil.C.Mg.ha_siteAverage-soil.C.Mg.ha_siteSD), ymax = (soil.C.Mg.ha_siteAverage+soil.C.Mg.ha_siteSD), colour = siteID), width = 0.2) +
        scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
        ggtitle("Soil Carbon Stocks by Site of D02") +
        ylab("Carbon Stock (Mg/ha)") +
        theme_classic(base_size = 16) +
        theme(plot.title = element_text(size=16, hjust = 0.5),
              axis.text.x=element_text(size=12, color = "black"),
              axis.text.y=element_text(size=12),
              axis.title.y=element_text(size=14), 
              axis.title.x=element_blank(), 
              legend.text=element_text(size=14),
              legend.title=element_text(size=14),
              legend.position = "none")


#Writing the joined dataframe into CSV file
write.csv(soil.by.plot.D02, "soil.by.plot.D02.csv")

```


#Nitrogen-----------------------------------------------------
##############################################################

```{r}
#install.packages("neonUtilities")						
 library(neonUtilities)								
 library(neonUtilities)					
 library(tidyverse)
 library(lubridate)
 library(dplyr)
 library(ggplot2)
 library(ggbreak)
 library(patchwork)
 
 #Necessary packages / commands for installling neonNTrans. Follow directions to update packages as needed. You can try to call for compilation for packages, but use binary if needed.
 #library(devtools) 
 #install_github("NEONScience/NEON-Nitrogen-Transformations/neonNTrans", #dependencies=TRUE)  
 #library(neonNTrans) 
 
```
 
 Downloading the data of interest from NEON, listing, and csvs
```{r}												
 
 #Commented out code is relevant if you are able to use neonNTrans package in your R version, which I (Anna) cannot
       # BLAN.soil.nitro <- loadByProduct(dpID="DP1.10086.001", 	
       # site=c("BLAN","SERC","SCBI"), 		
       # startdate="2012-01", enddate="2024-12", 				
       # package="basic")
       
       #list2env(BLAN.soil.nitro, .GlobalEnv)
       
       #spc_biogeochem
       #write.csv(spc_biogeochem, "spc_biogeochem.csv")
       
       #spc_bulkdensity
       #write.csv(spc_bulkdensity, "spc_bulkdensity.csv")
       
       
       # test.out <- def.calc.ntrans(kclInt = BLAN.soil.nitro$ntr_internalLab, kclIntBlank = BLAN.soil.nitro$ntr_internalLabBlanks, kclExt = BLAN.soil.nitro$ntr_externalLab, soilMoist = BLAN.soil.nitro$sls_soilMoisture, dropConditions = c("extract stored at incorrect temperature", "soil stored at incorrect temperature", "mass uncertain", "volume uncertain"))
 
 
data.summary.D02 <- read.csv("datasummary.D02.csv")
str(data.summary.D02)
 
 
 # Creating variable for siteID aand soil horizon s an independent variable in the dataframe by duplicating sample ID, subsetting the string, separating by the hyphen, and adding back to the original dataframe
data.summary.D02$siteID = substr(data.summary.D02$sampleID, 1,4)
data.summary.D02$siteID_tosplit = substr(data.summary.D02$sampleID, 1,10)
hold.name <- strsplit(data.summary.D02$siteID_tosplit, split="-")
hold.name.df <- as.data.frame(matrix(unlist(hold.name), ncol=2, byrow=T))
colnames(hold.name.df) <- c("plotID", "soilHorizon")
data.summary.D02$plotID <- hold.name.df$plotID
data.summary.D02$soilHorizon <- hold.name.df$soilHorizon
 
write.csv(data.summary.D02, "data.summary.D02_outputV2.csv")
 
```
 
 
 Notably, the Inorganic nitrogen metric provided through the neonNTrans package includes nitrite, which we are not as interested in, but some micorbes being able to utilize it as an intermediate. We could not a clearly identifiable way to isolate nitrate from nitrite with the available data.
 
 
# Figures at the plot-level by site
```{r}
 data.summary.D02.BLAN <- data.summary.D02 %>%
   filter(siteID == "BLAN")
 data.summary.D02.SERC <- data.summary.D02 %>%
   filter(siteID == "SERC")
 data.summary.D02.SCBI <- data.summary.D02 %>%
   filter(siteID == "SCBI")
 
 exploratory.InorgN.BLAN.plot <- ggplot(data=data.summary.D02.BLAN, aes(x=plotID, y=soilInorganicNugPerGram)) + 
   geom_point(size=1.5, alpha = 1)  +
         #scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
         ggtitle("Average Soil Inorganic N by Plot within BLAN of D02") +
        #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
         ylab("Inorganic N (mcg/g)") +
      geom_jitter() +
     theme_classic() +
   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
 exploratory.InorgN.BLAN.plot
 
 exploratory.InorgN.SERC.plot <- ggplot(data=data.summary.D02.SERC, aes(x=plotID, y=soilInorganicNugPerGram)) + 
   geom_point(size=1.5, alpha = 1)  +
         #scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
         ggtitle("Average Soil Inorganic N by Plot within SERC of D02") +
        #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
         ylab("Inorganic N (mcg/g)") +
      geom_jitter() +
     theme_classic() +
   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
 exploratory.InorgN.SERC.plot
 
 exploratory.InorgN.SCBIplot <- ggplot(data=data.summary.D02.SCBI, aes(x=plotID, y=soilInorganicNugPerGram)) + 
   geom_point(size=1.5, alpha = 1)  +
         #scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
         ggtitle("Average Soil Inorganic N by Plot within SCBI of D02") +
        #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
         ylab("Inorganic N (mcg/g)") +
      geom_jitter() +
     theme_classic() +
   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
 exploratory.InorgN.SCBIplot
 
 
```
 
# Figures at the site-level
```{r}
 
 # Initial visualization of data for Inorganic N
 data.summary.D02.Exploratory <- data.summary.D02 %>%
  group_by(siteID) %>%
  summarise(soilInorganicNugPerGram_average = mean(soilInorganicNugPerGram),
            soilInorganicNugPerGram_sd = sd(soilInorganicNugPerGram))
 
 exploratory.InorgN.D02 <- ggplot(data=data.summary.D02, aes(x=siteID, y=soilInorganicNugPerGram, group = siteID, color=siteID)) + 
   geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
   geom_point(data = data.summary.D02.Exploratory, aes(x = siteID, y = soilInorganicNugPerGram_average, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
   geom_errorbar(data = data.summary.D02.Exploratory, aes(x= siteID, y = soilInorganicNugPerGram_average, ymin = (soilInorganicNugPerGram_average-soilInorganicNugPerGram_sd), ymax = (soilInorganicNugPerGram_average+soilInorganicNugPerGram_sd), colour = siteID), width = 0.2) +
         scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
         ggtitle("Average Soil Inorganic N (ug/gram) by Site of D02") +
        scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
        #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
         ylab("Inorganic N (mcg/g)") +
         theme_classic(base_size = 16) +
         theme(plot.title = element_text(size=16, hjust = 0.5),
               axis.text.x=element_text(size=12, color = "black"),
               axis.text.y=element_text(size=12),
               axis.title.y=element_text(size=14), 
               axis.title.x=element_blank(), 
               legend.text=element_text(size=14),
               legend.title=element_text(size=14),
               legend.position = "none")
 
 # Initial visualization of data for Ammonium
 data.summary.D02.Exploratory.NH4 <- data.summary.D02 %>%
  group_by(siteID) %>%
  summarise(soilAmmoniumNugPerGram_average = mean(soilAmmoniumNugPerGram),
            soilAmmoniumNugPerGram_sd = sd(soilAmmoniumNugPerGram))
 
 exploratory.ammonium.D02 <- ggplot(data=data.summary.D02, aes(x=siteID, y=soilAmmoniumNugPerGram, group = siteID, color=siteID)) + 
   geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
   geom_point(data = data.summary.D02.Exploratory.NH4, aes(x = siteID, y = soilAmmoniumNugPerGram_average, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
   geom_errorbar(data = data.summary.D02.Exploratory.NH4, aes(x= siteID, y = soilAmmoniumNugPerGram_average, ymin = (soilAmmoniumNugPerGram_average-soilAmmoniumNugPerGram_sd), ymax = (soilAmmoniumNugPerGram_average+soilAmmoniumNugPerGram_sd), colour = siteID), width = 0.2) +
         scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
         ggtitle("Average Soil Ammonium (ug/gram) by Site of D02") +
        scale_y_continuous(breaks = seq(-10, 80, 10), limits = c(-10, 65)) +
        #scale_y_continuous(breaks = seq(-10, 80, 10), limits = c(-10, 65)) +
       scale_y_break(c(32, 48)) +
         ylab("Ammonium (mcg/g)") +
         theme_classic(base_size = 16) +
         theme(plot.title = element_text(size=16, hjust = 0.5),
               axis.text.x=element_text(size=12, color = "black"),
               axis.text.y=element_text(size=12),
               axis.title.y=element_text(size=14), 
               axis.title.x=element_blank(), 
               axis.text.y.right = element_blank(), #deals with weird ggbreak glitch
               axis.line.y.right = element_blank(), #deals with weird ggbreak glitch
               axis.ticks.y.right = element_blank(), #deals with weird ggbreak glitch
               legend.text=element_text(size=14),
               legend.title=element_text(size=14),
               legend.position = "none")
 
 exploratory.InorgN.D02
 exploratory.ammonium.D02
 
 # Initial visualization of data for Mineralization
 data.summary.D02.Exploratory.Mineralization <- data.summary.D02 %>%
  group_by(siteID) %>%
  summarise(netNminugPerGramPerDay_average = mean(netNminugPerGramPerDay), #averaging a rate, always messes with my head if appropriately statistically
            netNminugPerGramPerDay_sd = sd(netNminugPerGramPerDay))
 
 exploratory.mineralization.D02 <- ggplot(data=data.summary.D02, aes(x=siteID, y=netNminugPerGramPerDay, group = siteID, color=siteID)) + 
   geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
   geom_point(data = data.summary.D02.Exploratory.Mineralization, aes(x = siteID, y = netNminugPerGramPerDay_average, group = siteID, color=siteID), size = 3.5, alpha = 0.8, position = position_dodge(width = 0.5)) +
   geom_errorbar(data = data.summary.D02.Exploratory.Mineralization, aes(x= siteID, y = netNminugPerGramPerDay_average, ymin = (netNminugPerGramPerDay_average-netNminugPerGramPerDay_sd), ymax = (netNminugPerGramPerDay_average+netNminugPerGramPerDay_sd), colour = siteID), width = 0.2) +
         scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
         ggtitle("Average Soil N Mineralization rate ((ug/gram)/day) by Site of D02") +
        #scale_y_continuous(breaks = seq(-10, 80, 10), limits = c(-10, 65)) +
      # scale_y_break(c(32, 48)) +
         ylab("Mineralization ((mcg/g)/day)") +
         theme_classic(base_size = 16) +
         theme(plot.title = element_text(size=16, hjust = 0.5),
               axis.text.x=element_text(size=12, color = "black"),
               axis.text.y=element_text(size=12),
               axis.title.y=element_text(size=14), 
               axis.title.x=element_blank(), 
              #axis.text.y.right = element_blank(), #deals with weird ggbreak glitch
               #axis.line.y.right = element_blank(), #deals with weird ggbreak glitch
               #axis.ticks.y.right = element_blank(), #deals with weird ggbreak glitch
               legend.text=element_text(size=14),
               legend.title=element_text(size=14),
               legend.position = "none")
 
 exploratory.mineralization.D02 #note, getting a missing values warning
```
 
# Figures at the site-level
# Data Carpentry Figures at the site-level
```{r}
 
 D02.nitrogen.df <- read.csv("data.summary.D02_outputV2.csv")
 str(D02.nitrogen.df) #370 observations
 
 #When initially tried to join with carbon dataframe, was losing observations, likely because there are multiple observations by plot. Going to calculate by plot level for inorganic nitrogen and see if that was the issue...
 
 D02.nitrogen.df.plot <- D02.nitrogen.df %>%
   group_by(plotID) %>%
   summarise(soilInorganicNugPerGram.average = mean(soilInorganicNugPerGram)) #note, this is on a per gram basis, not on the plot scale... could calculate that...
 
 
 ##playing around calculating on the plot scale precisely
 
 #note this is contingent on having the other carbon data in environment
 str(soil.data.BLAN) #70
 str(soil.data.SERC) #103
 str(soil.data.SCBI) #80
 
 soil.data.D02.join1 <- full_join(soil.data.BLAN, soil.data.SERC)
 soil.data.D02.all <- full_join(soil.data.D02.join1, soil.data.SCBI)
 str(soil.data.D02.all) #looks like it adds up at 253
 #Okay, I want to extract bulk density averaged at the plot level to join onto to nitrogen dataframe...
 
 soil.data.D02.all.BDmeans.plots <- soil.data.D02.all %>%
   distinct(siteID, plotID, BD.plot.mean) %>%
   filter(!is.nan(BD.plot.mean))
 
 soil.data.D02.all.BDmeans.sites <- soil.data.D02.all %>%
   distinct(siteID, BD.site.mean) %>%
   group_by(siteID) %>%
   summarise(BD.site.mean = mean(BD.site.mean)) #had to do some coercive coding since unsure why distinct insisted on still grouping by plotID althought no specified
 
 #Now, to avoid excessive NAs, we are going to sequentially join to first add plot mean onto nitrogen dataframe, and then add on the site-level mean to ensure that doesn't have any NAs. From there, can then gap fill into a new bulk density variable to be able to calculate the total inorganic nitrogen that would be estimated at the plot level for each sample, to then use all those estimate to produce the 'average' estimate at the plot and site level for lm() with carbon, since they will now be on the same scale (I think)
 
 D02.nitrogen.df.BDinfo <- left_join(D02.nitrogen.df, soil.data.D02.all.BDmeans.plots)
 D02.nitrogen.df.BDinfo <- left_join(D02.nitrogen.df.BDinfo, soil.data.D02.all.BDmeans.sites) 
 
 #Appears to have been successful, now let's gap fill to make that bulk density variable in the dataframe
 #Telling R to see if the bulk density of the plot mean is available, and if it is not, then apply the value from the site mean variable, or else just keep the NA in the plot mean in this new variable
 D02.nitrogen.df.BDinfo <- D02.nitrogen.df.BDinfo %>%
   mutate(bulkdensity = if_else(is.na(BD.plot.mean), BD.site.mean, BD.plot.mean))
 
 sum(is.na(D02.nitrogen.df.BDinfo$bulkdensity)) #No NAs!
 
 
 #(1)	Given bulk density and the depth increment of each horizon in your dataframe, create a new column in your dataframe that is the total amount of soil in grams for a m^2 area
 #We used a m2 area, so use 100 cm x 100 cm  = 10,000 cm2 for the sample area in this case too.
 
 
D02.nitrogen.df.draft <- D02.nitrogen.df.BDinfo %>%   
mutate(sample.area.cm2 = 10000) %>%   
mutate(sample.volume = sample.area.cm2*30) %>% #30 cm is the max depth for what is in the user guide... but like, we can maybe figure out actualy depth from other tables output by package that I do not have?   
mutate(g.soil = sample.volume*bulkdensity) %>% #should give units of g/m2     
mutate(InorganicN.ugm2 = g.soil*soilInorganicNugPerGram) %>% # g/m2 * ug / g = ug/m2   
mutate(InorganicN.ugHectare = InorganicN.ugm2 * (10000/1)) %>% # ug/m2 * m2/ha = ug/ha   
mutate(InorganicN.gHectare = InorganicN.ugHectare * (1/1000000)) %>% #ug/ha * g/ug= g/ha   
mutate(InorganicN.kgHectare = InorganicN.gHectare * (1/1000)) # g/ha * kg/g = kg/ha 
 write.csv(D02.nitrogen.df.draft, "D02.nitrogen.df.draft.csv")
 
 
 # Initial visualization of data for Inorganic N
 D02.nitrogen.df.draft.Exploratory <- D02.nitrogen.df.draft %>%
  group_by(siteID, plotID) %>%
 summarise(InorganicN.kgHectare_average.plot = mean(InorganicN.kgHectare),
            InorganicN.kgHectare_sd.plot = sd(InorganicN.kgHectare))
 

 exploratory.InorgN.D02 <- ggplot(data=D02.nitrogen.df.draft.Exploratory, aes(x=siteID, y=InorganicN.kgHectare_average.plot, group = siteID, color=siteID)) + 
   geom_point(size=1.5, alpha = 0.3, position = position_jitterdodge(dodge.width = 0.5))  +
         scale_color_brewer(name = "SiteID", palette = "Dark2") + #color-blind friendly
         ggtitle("Average Soil Inorganic N per 1 Hectare by Site of D02") +
        #scale_y_continuous(breaks = seq(0, 100, 10), limits = c(-15, 100)) +
         ylab("Inorganic N (kg/hectare)") +
         theme_classic(base_size = 16) +
         theme(plot.title = element_text(size=16, hjust = 0.5),
               axis.text.x=element_text(size=12, color = "black"),
               axis.text.y=element_text(size=12),
               axis.title.y=element_text(size=14), 
               axis.title.x=element_blank(), 
               legend.text=element_text(size=14),
               legend.title=element_text(size=14),
               legend.position = "none")
 exploratory.InorgN.D02
 
 
```

```{r}
D02.carbon.merged.step1.df <- full_join(AGB.trees.by.plot.D02, roots.by.plot.D02)
D02.carbon.merged.df <- full_join(D02.carbon.merged.step1.df,soil.by.plot.D02)

write.csv(D02.carbon.merged.df, "D02.carbon.merged.csv")
```

#Modeling-------------------------------------------------------------------------
####Not sure I'm doing this right/using the right code.####




# Soil ##########################

```{r}
soil.model.test <- lme(InorganicN.kgHectare ~ soil.C.Mg.ha + domainID + soil.C.Mg.ha*domainID,
                    random = ~1 | siteID/plotID,
                    correlation = corAR1(), #currently left blank for now, it seems like it'll just go to default value, but looks like it may be good practice to specify the time variable and nesting structure here. Unsure if that would appropriately set autocorrelation related to the inorganic N
                    data = D02_D03_D08_D19_fulljoin_CN_test,
                    na.action=na.exclude)
summary(soil.model.test)
# No significant main effect of soil carbon
# Significant interaction between soil carbon and D03, when compared to D02 as the reference domain. Specifically, the effect of soil carbon on inorganic N is modified by +0.243 units compared to the reference domain

```


# Tree ##########################

```{r}
tree.model.test <- lme(InorganicN.kgHectare ~ AGB.C.Mg.ha + domainID + AGB.C.Mg.ha*domainID,
                    random = ~1 | siteID/plotID,
                    correlation = corAR1(), #currently left blank for now, it seems like it'll just go to default value, but looks like it may be good practice to specify the time variable and nesting structure here. Unsure if that would appropriately set autocorrelation related to the inorganic N
                    data = D02_D03_D08_D19_fulljoin_CN_test,
                    na.action=na.exclude)
summary(tree.model.test)
# No significant main effect of soil carbon
# Significant interaction between soil carbon and D19, when compared to D02 as the reference domain. Specifically, the effect of soil carbon on inorganic N is modified by +0.236 units compared to the reference domain
```


# Roots ##########################

```{r}
roots.model.test <- lme(InorganicN.kgHectare ~ mean.root.MgC.ha + domainID + mean.root.MgC.ha*domainID,
                    random = ~1 | siteID/plotID,
                    correlation = corAR1(), #currently left blank for now, it seems like it'll just go to default value, but looks like it may be good practice to specify the time variable and nesting structure here. Unsure if that would appropriately set autocorrelation related to the inorganic N
                    data = D02_D03_D08_D19_fulljoin_CN_test,
                    na.action=na.exclude)
summary(roots.model.test)
# Significant main effect of soil carbon
# Significant main effect of D03 relative to D02
# Significant main effect of D08 relative to D02
# Significant interaction between root carbon and D03
# Significant interaction between root carbon and D08
# Significant interaction between root carbon and D19
```


#Undecided------------------------------------------------------------------------

#Working Code, still need to add Nitrogen for D19--- Joining C and N dataframes for D03; Carpentry to join by domain
```{r}

#Domain split for tracking self
# D03 = Anna
# D02 = Andrea
# D19 = Kylie
# D08 = Trey

# General Joining Approach -- Join carbon and nitrogen by domain initially; Then join these dataframes across domains. 

# D03 #######################################
##(1) Adding in Domain ID for each dataframe
D03.carbon.merged.df <- D03.carbon.merged.df %>%
  mutate(domainID = "D03")

D03.nitrogen.df.draft <- D03.nitrogen.df.draft %>%
    mutate(domainID = "D03")
  
##(2) Selecting variables needed to streamline bulkiness of df
D03.streamlined.carbon.df <- D03.carbon.merged.df %>%
  select(domainID, siteID, plotID, AGB.C.Mg.ha, mean.root.MgC.ha, soil.C.Mg.ha, totalSampledAreaTrees, plot.AGB)

D03.streamlined.nitrogen.df <- D03.nitrogen.df.draft %>%
  select(domainID, siteID, plotID, sampleID, collectDate, InorganicN.kgHectare, soilAmmoniumNugPerGram, soilInorganicNugPerGram, netNminugPerGramPerDay, netNitugPerGramPerDay)

##(3) Joining the streamlined carbon and nitrogen dataframes together for D02
D03.fulljoin.CN.test.df <- full_join(D03.streamlined.carbon.df, D03.streamlined.nitrogen.df)
## str(D03.fulljoin.CN.test.df) # 481 x 15


#D02 #######################################
## Read in the data
D02.carbon.merged.df <- read.csv("D02.carbon.merged.csv")
# str(D02.carbon.merged.df) #136 observations

D02.nitrogen.merged.df <- read.csv("D02.nitrogen.df.draft.csv")
# str(D02.nitrogen.merged.df)  #345 observations


##(1) Adding in Domain ID
D02.carbon.merged.df <- D02.carbon.merged.df %>%
    mutate(domainID = "D02")

D02.nitrogen.merged.df <- D02.nitrogen.merged.df %>%
    mutate(domainID = "D02")

##(2) Selecting variables needed to streamline bulkiness of df
D02.streamlined.carbon.df <- D02.carbon.merged.df %>%
  select(domainID, siteID, plotID, AGB.C.Mg.ha, mean.root.MgC.ha, soil.C.Mg.ha, totalSampledAreaTrees, plot.AGB)
# str(D02.streamlined.carbon.df)

D02.streamlined.nitrogen.df <- D02.nitrogen.merged.df %>%
  select(domainID, siteID, plotID, sampleID, collectDate, InorganicN.kgHectare, soilAmmoniumNugPerGram, soilInorganicNugPerGram, netNminugPerGramPerDay, netNitugPerGramPerDay)
# str(D02.streamlined.nitrogen.df)

##(3) Joining the streamlined carbon and nitrogen dataframes together for D02
D02.fulljoin.CN.test.df <- full_join(D02.streamlined.carbon.df, D02.streamlined.nitrogen.df)
## str(D02.fulljoin.CN.test.df) # 481 x 15


#D19 #######################################

# Commenting out this code when initially trying to address merging issues with original merged data, went back and reconstructed merged carbon dataframe #######

# Reading in the merged carbon data for D19
# D19.carbon.merged.df <- read.csv("D19.carbon.merged.csv")
# # str(D19.carbon.merged.df)
# 
# ###Added Data carpentry to try to fix the many to many relationship
# D19.carbon.merged.df$siteID.corrected = substr(D19.carbon.merged.df$plotID, 1,4) #issue with NAs being present for site ID propogating error in merging
# D19.carbon.merged.df <- D19.carbon.merged.df %>%
#   filter(siteID.corrected == "BONA" | siteID.corrected == "DEJU" | siteID.corrected == "HEAL") #Removing megapit values, check with Kylie and see if that makes sense
# D19.carbon.merged.df <- D19.carbon.merged.df %>%
#   select(-c(siteID)) #removing original siteID variable with NAs
# D19.carbon.merged.df <- D19.carbon.merged.df %>% 
#   rename(siteID = siteID.corrected) #renaming the corrected siteID column to just "siteID" for merging purposes

# Reading in the merged nitrogen data for D19
D19.nitrogen.merged.df <- read.csv("D19.nitrogen.df.draft.csv") #There are NAs in the plotID column making, seems like it is causing an issue to merge; Seems like they have siteID and a mean bulk density for the plot but not plotID... will remove for now to see if helps joining issue

D19.nitrogen.merged.df <- D19.nitrogen.merged.df %>%
  filter(!is.na(plotID)) #removing NAs in plotID
# str(D19.nitrogen.merged.df)


#Carbon dataframe -- Trying to fix merging issues
AGB.trees.by.plot.D19 <- read.csv("AGB.trees.by.plot.D19.csv")
roots.by.plot.D19 <- read.csv("roots.by.plot.D019.csv") ## Backtraced source of error for OSBS in D19... Not sure why/how that got in there, can fix on the back end because confirmed these sites for OSBS are accounted for in D03
roots.by.plot.D19 <- roots.by.plot.D19 %>%
  filter(siteID == "BONA" | siteID == "DEJU") #Note, there is no HEAL for D19 for roots.

soil.by.plot.D19 <- read.csv("soil.by.plot.D19.csv")

D19.carbon.merged.step1.df <- full_join(AGB.trees.by.plot.D19, roots.by.plot.D19, by=join_by(siteID, plotID)) 
D19.carbon.merged.Anna.df <- full_join(D19.carbon.merged.step1.df, soil.by.plot.D19, by=join_by(siteID, plotID))
# str(D08.carbon.merged.df) #131
D19.carbon.merged.Anna.df <- D19.carbon.merged.Anna.df[-c(77, 78), ] #Removing megapit values, check with Kylie and see if that makes sense
write.csv(D19.carbon.merged.Anna.df, "D19.carbon.merged.Anna.csv")


##(1) Adding in Domain ID
D19.carbon.merged.df <- D19.carbon.merged.Anna.df %>%
  mutate(domainID = "D19")

D19.nitrogen.merged.df <- D19.nitrogen.merged.df %>%
  mutate(domainID = "D19")


##(2) Selecting variables needed to streamline bulkiness of df
D19.streamlined.carbon.df <- D19.carbon.merged.df %>%
  select(domainID, siteID, plotID, AGB.C.Mg.ha, mean.root.MgC.ha, soil.C.Mg.ha, totalSampledAreaTrees, plot.AGB)
 # str(D19.streamlined.carbon.df) # 95 x 8

D19.streamlined.nitrogen.df <- D19.nitrogen.merged.df %>%
  select(domainID, siteID, plotID, sampleID, collectDate, InorganicN.kgHectare, soilAmmoniumNugPerGram, soilInorganicNugPerGram, netNminugPerGramPerDay, netNitugPerGramPerDay)
# str(D19.streamlined.nitrogen.df)

##(3) Joining the streamlined carbon and nitrogen dataframes together for D02
D19.fulljoin.CN.test.df <- full_join(D19.streamlined.carbon.df, D19.streamlined.nitrogen.df) #getting a many-to-many relationship warning; iteratively going, seems like the many-to-many is related to DEJU_045.... Went ahead and remerged carbon dataframe as that seemed to be the source of the issue; Doing so seemed to fix the warning and visual inspection looks correct


#D08 ##################################
## Will need to merge carbon dataframe
AGB.trees.by.plot.D08 <- read.csv("AGB.trees.by.plot.D08.csv")
roots.by.plot.D08 <- read.csv("roots.by.plot.D08.csv")
soil.by.plot.D08 <- read.csv("soil.by.plot.D08.csv")

D08.carbon.merged.step1.df <- full_join(AGB.trees.by.plot.D08, roots.by.plot.D08, by=join_by(siteID, plotID)) 
D08.carbon.merged.df <- full_join(D08.carbon.merged.step1.df, soil.by.plot.D08, by=join_by(siteID, plotID))
# str(D08.carbon.merged.df) #131

write.csv(D08.carbon.merged.df, "D08.carbon.merged.csv")

## Reading in the merged carbon data for D08
D08.carbon.merged.df <- read.csv("D08.carbon.merged.csv")
# str(D08.carbon.merged.df) #131

D08.nitrogen.merged.df <- read.csv("D08.nitrogen.df.draft.csv")
# str(D08.nitrogen.merged.df)  #443 observations

##(1) Adding in Domain ID
D08.carbon.merged.df <- D08.carbon.merged.df %>%
    mutate(domainID = "D08")

D08.nitrogen.merged.df <- D08.nitrogen.merged.df %>%
    mutate(domainID = "D08")

##(2) Selecting variables needed to streamline bulkiness of df
D08.streamlined.carbon.df <- D08.carbon.merged.df %>%
  select(domainID, siteID, plotID, AGB.C.Mg.ha, mean.root.MgC.ha, soil.C.Mg.ha, totalSampledAreaTrees, plot.AGB)
# str(D08.streamlined.carbon.df) #224

D08.streamlined.nitrogen.df <- D08.nitrogen.merged.df %>%
  select(domainID, siteID, plotID, sampleID, collectDate, InorganicN.kgHectare, soilAmmoniumNugPerGram, soilInorganicNugPerGram, netNminugPerGramPerDay, netNitugPerGramPerDay)
# str(D08.streamlined.nitrogen.df) #443


##(3) Joining the streamlined carbon and nitrogen dataframes together for D02
D08.fulljoin.CN.test.df <- full_join(D08.streamlined.carbon.df, D08.streamlined.nitrogen.df) 
#Note, initially had a many-to-many error, and seems like its coming from plotID not appropriately merging at the initial joins. Going back to check if specifying only plotID and siteID fix it, forcing it to ignore 'X' variable... Did that specification for lines 1500-1501. After inspecting, looks like that did the trick.

# str(D08.fulljoin.CN.test.df) # 554 x 15


#Testing the join of D03, D02, D08 ##################################
D02.D03.fulljoin.CN.test.df <- full_join(D02.fulljoin.CN.test.df, D03.fulljoin.CN.test.df)
D02.D03.D08.fulljoin.CN.test.df <- full_join(D02.D03.fulljoin.CN.test.df, D08.fulljoin.CN.test.df)
D02.D03.D08.D19.fulljoin.CN.test.df <- full_join(D02.D03.D08.fulljoin.CN.test.df, D19.fulljoin.CN.test.df)
# str(D02.D03.D08.D19.fulljoin.CN.test.df) #1763

write.csv(D02.D03.D08.D19.fulljoin.CN.test.df, "D02.D03.D08.D19.fulljoin.CN.test.csv")


# Initial created for autocorrelation field, doesn't seem to be needed though so commented out
# D02.D03.D08.fulljoin.CN.test.df$year= substr(D02.D03.D08.fulljoin.CN.test.df$collectDate, 1,4)
  


D02.D03.D08.D19.fulljoin.CN.test.df %>%
  select(domainID, siteID) %>%
  distinct() %>%  # ensure uniqueness
  count(domainID, name = "n_sites") %>%
  arrange(desc(n_sites))  # I do not know why D19 has four sites... When did OSBS get in there?

```


#Code Under Construction, model building -- using lme() in nlme package (main advantage is can utilize autocorrelation term)
```{r}
#Ensuring variables are treated as factors
D02_D03_D08_D19_fulljoin_CN_test$domainID <- as.factor(D02_D03_D08_D19_fulljoin_CN_test$domainID)
D02_D03_D08_D19_fulljoin_CN_test$siteID <- as.factor(D02_D03_D08_D19_fulljoin_CN_test$siteID)

# Soil ##########################
soil.model.test <- lme(InorganicN.kgHectare ~ soil.C.Mg.ha + domainID + soil.C.Mg.ha*domainID,
                    random = ~1 | siteID/plotID,
                    correlation = corAR1(), #currently left blank for now, it seems like it'll just go to default value, but looks like it may be good practice to specify the time variable and nesting structure here. Unsure if that would appropriately set autocorrelation related to the inorganic N
                    data = D02_D03_D08_D19_fulljoin_CN_test,
                    na.action=na.exclude)
summary(soil.model.test)
# No significant main effect of soil carbon
# Significant interaction between soil carbon and D03, when compared to D02 as the reference domain. Specifically, the effect of soil carbon on inorganic N is modified by +0.243 units compared to the reference domain

# Tree ##########################
tree.model.test <- lme(InorganicN.kgHectare ~ AGB.C.Mg.ha + domainID + AGB.C.Mg.ha*domainID,
                    random = ~1 | siteID/plotID,
                    correlation = corAR1(), #currently left blank for now, it seems like it'll just go to default value, but looks like it may be good practice to specify the time variable and nesting structure here. Unsure if that would appropriately set autocorrelation related to the inorganic N
                    data = D02_D03_D08_D19_fulljoin_CN_test,
                    na.action=na.exclude)
summary(tree.model.test)
# No significant main effect of soil carbon
# Significant interaction between soil carbon and D19, when compared to D02 as the reference domain. Specifically, the effect of soil carbon on inorganic N is modified by +0.236 units compared to the reference domain

# Roots ##########################
roots.model.test <- lme(InorganicN.kgHectare ~ mean.root.MgC.ha + domainID + mean.root.MgC.ha*domainID,
                    random = ~1 | siteID/plotID,
                    correlation = corAR1(), #currently left blank for now, it seems like it'll just go to default value, but looks like it may be good practice to specify the time variable and nesting structure here. Unsure if that would appropriately set autocorrelation related to the inorganic N
                    data = D02_D03_D08_D19_fulljoin_CN_test,
                    na.action=na.exclude)
summary(roots.model.test)
# Significant main effect of soil carbon
# Significant main effect of D03 relative to D02
# Significant main effect of D08 relative to D02
# Significant interaction between root carbon and D03
# Significant interaction between root carbon and D08
# Significant interaction between root carbon and D19

```
 
#Code Under Construction, model visualization -- based on lme() in nlme package
```{R}
###### Preliminary model visualization, will likely need to split out due to y-axis by domain for appropriate resolution; Can also use min(, na.rm=TRUE) and max(, na.rm=TRUE) to identify the extreme variables skewing the axis

#Learning that the predict function I was trying to use to extract the predicted values from the nlme model does not tolerate NAs, which from my understanding automatically get excluded when running the nlme model. So, needed to create a new subdataframe without NAs in the variables of our model in order to graph those predicted lines.


# Soil ###########################
#(1) Create subset dataframe without NAs in model variable for graphing purposes using predict
soil_predict_df <- D02_D03_D08_D19_fulljoin_CN_test %>%
  filter(!is.na(soil.C.Mg.ha), !is.na(InorganicN.kgHectare), !is.na(siteID))
soil_predict_df$pred <- predict(soil.model.test, newdata = soil_predict_df, level=1)

#(2) plot the raw data and the predicted values for the model
soil_plot <- ggplot(D02_D03_D08_D19_fulljoin_CN_test, aes(x = soil.C.Mg.ha, y = InorganicN.kgHectare, colour = siteID)) +
      facet_wrap(~domainID, nrow=2) +   # a panel for each mountain range
      geom_point(alpha = 0.5) +
      theme_classic() +
     geom_line(data = soil_predict_df, aes(x = soil.C.Mg.ha, y = pred, group = siteID), linewidth = 1) + # adding predicted line from mixed model 
     #scale_y_continuous(breaks = seq(0, 75, 25), limits = c(-10, 80)) +
      theme(panel.spacing = unit(2, "lines"))  # adding space between panels
soil_plot #note the warning message, pretty sure these are related to the intended empty rows via the NAs

# Trees ###########################
#(1) Create subset dataframe without NAs in model variable for graphing purposes using predict
tree_predict_df <- D02_D03_D08_D19_fulljoin_CN_test %>%
  filter(!is.na(AGB.C.Mg.ha), !is.na(InorganicN.kgHectare), !is.na(siteID))
tree_predict_df$pred <- predict(tree.model.test, newdata = tree_predict_df, level=1)

#(2) plot the raw data and the predicted values for the model
tree_plot <- ggplot(D02_D03_D08_D19_fulljoin_CN_test, aes(x = AGB.C.Mg.ha, y = InorganicN.kgHectare, colour = siteID)) +
      facet_wrap(~domainID, nrow=2) +   # a panel for each mountain range
      geom_point(alpha = 0.5) +
      theme_classic() +
     geom_line(data = tree_predict_df, aes(x = AGB.C.Mg.ha, y = pred, group = siteID), linewidth = 1) + # adding predicted line from mixed model 
     scale_y_continuous(breaks = seq(0, 75, 25), limits = c(-10, 80)) +
      theme(panel.spacing = unit(2, "lines"))  # adding space between panels
tree_plot #note the warning message, pretty sure these are related to the intended empty rows via the NAs



# Roots ###########################
#(1) Create subset dataframe without NAs in model variable for graphing purposes using predict
root_predict_df <- D02_D03_D08_D19_fulljoin_CN_test %>%
  filter(!is.na(mean.root.MgC.ha), !is.na(InorganicN.kgHectare), !is.na(siteID))
root_predict_df$pred <- predict(roots.model.test, newdata = root_predict_df, level=1)

#(2) plot the raw data and the predicted values for the model
root_plot <- ggplot(D02_D03_D08_D19_fulljoin_CN_test, aes(x = mean.root.MgC.ha, y = InorganicN.kgHectare, colour = siteID)) +
      facet_wrap(~domainID, nrow=2) +   # a panel for each mountain range
      geom_point(alpha = 0.5) +
      theme_classic() +
    scale_y_continuous(breaks = seq(0, 75, 25), limits = c(-10, 80)) + #note, this does exclude one point from BLAN at 150, be mindful
     geom_line(data = root_predict_df, aes(x = mean.root.MgC.ha, y = pred, group = siteID), linewidth = 1) + # adding predicted line from mixed model 
    theme(panel.spacing = unit(2, "lines"))  # adding space between panels
root_plot

```



# Post-hoc using emmtrends
```{r}
library(emmeans) #emmtrends in emmeans package automatically detects nlme structure and even the random effects

#For models with interactions, we need to do pairwise comparisons to understand how the relationship between carbon stock and inorganic N stock varies by domain

#Soil
soil.emtrends <- emtrends(soil.model.test, specs = "domainID", var = "soil.C.Mg.ha")
pairs(soil.emtrends)
soil.trends.df <- as.data.frame(soil.emtrends)
soil.emtrends.plot <- ggplot(soil.trends.df, aes(x = domainID, y = soil.C.Mg.ha.trend)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  labs(title = "Estimated Slope of Soil Carbon by Domain",
    y = "Slope of Inorganic N ~ Soil Carbon Stock",
    x = "Domain") +
  theme_classic()
soil.emtrends.plot

#Tree
tree.emtrends <- emtrends(tree.model.test, specs = "domainID", var = "AGB.C.Mg.ha")
pairs(tree.emtrends)
tree.trends.df <- as.data.frame(tree.emtrends)
tree.emtrends.plot <- ggplot(tree.trends.df, aes(x = domainID, y = AGB.C.Mg.ha.trend)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  labs(title = "Estimated Slope of Tree Carbon by Domain",
    y = "Slope of Inorganic N ~ Tree Carbon Stock",
    x = "Domain") +
  theme_classic()
tree.emtrends.plot

#Roots
pairs(emtrends(roots.model.test, specs = "domainID", var = "mean.root.MgC.ha")) 
    #double check how many sites are included for root carbon stock for D19...
    print.d19 <- D02_D03_D08_D19_fulljoin_CN_test %>%
      filter(domainID == "D19")
    unique(print.d19$siteID) #Why is OSBS a part of D19? Backtraced error up to import of roots.by.plot.D19. Original dataframe incorrect and had OSBS in it.
    
roots.emtrends <- emtrends(roots.model.test, specs = "domainID", var = "mean.root.MgC.ha")
pairs(roots.emtrends)
roots.trends.df <- as.data.frame(roots.emtrends)
roots.emtrends.plot <- ggplot(roots.trends.df, aes(x = domainID, y = mean.root.MgC.ha.trend)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  labs(title = "Estimated Slope of roots Carbon by Domain",
    y = "Slope of Inorganic N ~ roots Carbon Stock",
    x = "Domain") +
  theme_classic()
roots.emtrends.plot

```










